---
title: "Chapter 3: Plasticity of functional traits and its evolutionary constraints in *Nemophila menziesii*"
author: "Helen E. Payne"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: false
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

This R Markdown reproduces the analyses reported in *Chapter 3:
Plasticity of functional traits and its evolutionary constraints in*
Nemophila menziesii. Section numbers follow the Methods/Results
structure of the manuscript; equations (Eq. 1–8) refer to the manuscript
equations. Four pedigreed populations are analysed in parallel: Angelo
Coast (AC), Bodega Bay (BB), Blue Oak (BO), and Hastings Reserve (HR).

Plasticity–fitness analyses (Eqs. 5–8) are run twice:

* **Primary** — fitness `W` is the aster sire effect on the mean-value
  scale (BLUP from `chapter_1_adaptive_capacity/aster/Aster6_*.Rmd`);
  this is the metric used in Chapter 1 and integrates the full
  multi-stage life history.
* **Sensitivity** — fitness `W` is the family-mean observed fecundity
  (`est_fecundity = total_fruits × mean_seeds_per_fruit`); this matches
  the convention of Chapter 2 and avoids regression dilution from BLUP
  shrinkage.

A side-by-side comparison appears in §11. The qualitative conclusion is
robust to the choice of fitness metric, but the specific trait ×
population combinations that reach p < 0.05 differ between the two,
which we interpret as evidence that isolated significant associations
likely reflect false positives under the multiple-testing burden of
4 traits × 4 sites × multiple model frameworks.

# Setup

```{r libraries}
library(tidyverse)
library(lme4)
library(lmerTest)
library(performance)
library(broom)
library(broom.mixed)
library(patchwork)
library(here)
library(knitr)
library(kableExtra)
```

```{r constants}
SITES  <- c("AC", "BB", "BO", "HR")
TRAITS <- c("corolla_diam_mm_SEG", "SLA_SEG", "d13C_SEG", "msm_all")
# PIv (Eq. 1) is undefined for traits on a signed/negative scale (δ13C ≈ -25‰);
# normalization by |max| produces scale-dependent values that are not
# comparable across populations. Directional plasticity (Eq. 2) is used for
# δ13C instead.
PIv_TRAITS <- setdiff(TRAITS, "d13C_SEG")
TRAIT_LABELS <- c(
  corolla_diam_mm_SEG = "corolla diameter (mm)",
  SLA_SEG             = "specific leaf area (mm²/mg)",
  d13C_SEG            = "δ13C (‰)",
  msm_all             = "mean seed mass (mg)"
)
```

# 1. Data loading and trait preparation

The 2022 compiledsheet contains only Generation 1 plants; in 2023 we
restrict to `Gen == "G1"` so the comparison matches Year 1. We retain
only designated fitness plants (`F_plant == TRUE`). Each row carries
both the four focal trait values and `est_fecundity`, the per-plant
observed fecundity used in the sensitivity analysis.

```{r load-site}
load_site <- function(site) {
  read_one <- function(year_int, gen_filter) {
    path <- here("data_sheets/compiled_sheets",
                 paste0(site, "_compiledsheet_full_", year_int, ".csv"))
    d <- read_csv(path, show_col_types = FALSE)
    if (gen_filter) d <- d %>% filter(Gen == "G1")
    d %>%
      mutate(
        # Chapter 2 survival correction: a plant that produced fruits must
        # have survived to flowering.
        surv_to_flower = ifelse(total_fruits > 0 & surv_to_flower == 0,
                                1, surv_to_flower),
        # Plants that did not survive (or were unmeasured) contribute zero
        # fecundity, not NA.
        est_fecundity  = ifelse(is.na(est_fecundity) | surv_to_flower == 0,
                                0, est_fecundity)
      ) %>%
      filter(F_plant == TRUE) %>%
      dplyr::select(Donor, Recipient, all_of(TRAITS), est_fecundity) %>%
      mutate(Year = as.character(year_int), Site = site)
  }
  bind_rows(read_one(2022, FALSE), read_one(2023, TRUE)) %>%
    mutate(Site = factor(Site, levels = SITES),
           Year = factor(Year, levels = c("2022", "2023")))
}

site_data <- set_names(purrr::map(SITES, load_site), SITES)
walk2(SITES, site_data, ~ cat(.x, ":", nrow(.y), "rows\n"))
```

# 2. Trait-specific plasticity (PIv, PImd, directional)

Per-family per-trait means are the foundation for every plasticity
metric. PIv (Eq. 1) is the relative range of family means across the
two years. PImd is the median-based analogue (used as a robustness
check). Directional plasticity (Eq. 2) is the signed proportional change
from Year 1 to Year 2.

```{r plasticity-fns}
family_means <- function(d) {
  d %>%
    group_by(Donor, Year) %>%
    summarise(across(all_of(TRAITS), ~ mean(.x, na.rm = TRUE),
                     .names = "{.col}"),
              .groups = "drop") %>%
    pivot_longer(all_of(TRAITS), names_to = "Trait", values_to = "mean")
}

family_medians <- function(d) {
  d %>%
    pivot_longer(all_of(TRAITS), names_to = "Trait", values_to = "value") %>%
    group_by(Donor, Year, Trait) %>%
    summarise(median = median(value, na.rm = TRUE), .groups = "drop")
}

compute_PIv <- function(fam_means) {           # Eq. 1
  fam_means %>%
    group_by(Donor, Trait) %>%
    summarise(max_mean = max(mean, na.rm = TRUE),
              min_mean = min(mean, na.rm = TRUE),
              PIv      = (max_mean - min_mean) / abs(max_mean),
              .groups  = "drop")
}

compute_PImd <- function(fam_medians) {
  fam_medians %>%
    group_by(Donor, Trait) %>%
    summarise(PImd = (max(median, na.rm = TRUE) - min(median, na.rm = TRUE)) /
                       abs(max(median, na.rm = TRUE)),
              .groups = "drop")
}

compute_directional <- function(fam_means) {   # Eq. 2
  fam_means %>%
    pivot_wider(names_from = Year, values_from = mean,
                names_prefix = "Y") %>%
    mutate(directional = (Y2023 - Y2022) / abs(Y2022)) %>%
    dplyr::select(Donor, Trait, directional, Y2022, Y2023)
}

plasticity <- purrr::map(site_data, function(d) {
  fm   <- family_means(d)
  fmed <- family_medians(d)
  list(
    fam_means    = fm,
    PIv          = compute_PIv(fm),
    PImd         = compute_PImd(fmed),
    directional  = compute_directional(fm)
  )
})

# Mean directional plasticity across all 4 traits per family (multivariate
# directional), used in Figure 8 and Eqs 5/6 multivariate panels.
mean_directional_by_family <- purrr::map_dfr(SITES, function(site) {
  plasticity[[site]]$directional %>%
    group_by(Donor) %>%
    summarise(mean_directional = mean(directional, na.rm = TRUE),
              .groups = "drop") %>%
    mutate(Site = site, .before = 1)
})
```

## PIv vs PImd robustness check (Table S4)

```{r piv-pimd-cor}
piv_pimd_cor <- purrr::map_dfr(SITES, function(site) {
  x <- plasticity[[site]]
  inner_join(x$PIv  %>% filter(Trait %in% PIv_TRAITS),
             x$PImd %>% filter(Trait %in% PIv_TRAITS),
             by = c("Donor", "Trait")) %>%
    group_by(Trait) %>%
    summarise(n   = n(),
              r   = cor(PIv, PImd, use = "complete.obs"),
              p   = cor.test(PIv, PImd)$p.value,
              .groups = "drop") %>%
    mutate(Site = site, .before = 1)
})
piv_pimd_cor
```

# 3. Comparing trait-specific plasticity (Table S6)

Trait-level mean PIv (across paternal families) summarises how plastic
each trait is on average in each population. The one-sample t-test
evaluates whether PIv differs from zero.

```{r piv-by-trait}
piv_by_trait <- purrr::map_dfr(SITES, function(site) {
  plasticity[[site]]$PIv %>%
    filter(Trait %in% PIv_TRAITS) %>%
    group_by(Trait) %>%
    summarise(n        = n(),
              mean_PIv = mean(PIv, na.rm = TRUE),
              sd_PIv   = sd(PIv,   na.rm = TRUE),
              se_PIv   = sd_PIv / sqrt(n),
              t        = t.test(PIv, mu = 0)$statistic,
              p        = t.test(PIv, mu = 0)$p.value,
              .groups  = "drop") %>%
    mutate(Site = site, .before = 1)
})
piv_by_trait
```

# 4. Year effects on mean trait values (Eq. 3)

For each trait × site we fit `trait ~ Year + (1|Donor) + (1|Recipient)`,
report Year fixed-effect estimates and marginal/conditional R².

```{r year-lmm}
fit_year_lmm <- function(d, trait) {
  f <- as.formula(paste0(trait, " ~ Year + (1|Donor) + (1|Recipient)"))
  lmer(f, data = d, REML = FALSE)
}

year_lmm_results <- purrr::map_dfr(SITES, function(site) {
  purrr::map_dfr(TRAITS, function(tr) {
    m <- fit_year_lmm(site_data[[site]], tr)
    r <- suppressWarnings(performance::r2(m))
    fe <- broom.mixed::tidy(m, effects = "fixed") %>%
      filter(term == "Year2023")
    tibble(Site = site, Trait = tr,
           estimate = fe$estimate, std.error = fe$std.error,
           p_Year   = fe$p.value,
           R2m      = as.numeric(r$R2_marginal),
           R2c      = as.numeric(r$R2_conditional),
           n        = nobs(m))
  })
})
year_lmm_results
```

# 5. Genotype × Year interactions (Eq. 4, Figure 7)

The null model is Eq. 3; the full model adds `(1|Donor:Year)`. The
significance of the `Donor:Year` variance component is assessed with a
REML likelihood ratio test; the p-value is halved to correct for testing
on the boundary (H₀: σ² = 0; Self & Liang 1987).  Both models are fit
with REML; variance components are drawn from the REML full model.

```{r gxy-lmm}
fit_gxy <- function(d, trait) {
  fn <- as.formula(paste0(trait, " ~ Year + (1|Donor) + (1|Recipient)"))
  fg <- as.formula(paste0(trait,
                          " ~ Year + (1|Donor) + (1|Recipient) + (1|Donor:Year)"))
  null <- lmer(fn, data = d, REML = TRUE)
  full <- lmer(fg, data = d, REML = TRUE)
  list(null = null, full = full,
       lrt_reml = anova(null, full, refit = FALSE))
}

gxy_models <- purrr::map(SITES, function(site) {
  set_names(purrr::map(TRAITS, ~ fit_gxy(site_data[[site]], .x)), TRAITS)
}) %>% set_names(SITES)

# REML LRT summary with boundary-corrected p (halved)
gxy_lrt <- purrr::map_dfr(SITES, function(site) {
  purrr::map_dfr(TRAITS, function(tr) {
    a <- gxy_models[[site]][[tr]]$lrt_reml
    tibble(Site = site, Trait = tr,
           chisq = a$Chisq[2],
           p_boundary = a$`Pr(>Chisq)`[2] / 2)
  })
})
gxy_lrt

# Variance components as % of total variance (input to Figure 7)
gxy_varcomp <- purrr::map_dfr(SITES, function(site) {
  purrr::map_dfr(TRAITS, function(tr) {
    vc <- as.data.frame(VarCorr(gxy_models[[site]][[tr]]$full))
    vc$percent <- 100 * vc$vcov / sum(vc$vcov)
    vc %>% mutate(Site = site, Trait = tr, .before = 1) %>%
      dplyr::select(Site, Trait, grp, vcov, sdcor, percent)
  })
})
```

## Variance components table

Each cell shows `σ² (±SD, % of total variance)` for the four random
components of the G×Y full model (Eq. 4).  `n` is the number of
observations contributing to each model fit.

```{r gxy-varcomp-table}
TRAIT_LABELS_SHORT <- c(
  corolla_diam_mm_SEG = "Corolla diameter",
  SLA_SEG             = "SLA",
  d13C_SEG            = "δ¹³C",
  msm_all             = "Mean seed mass"
)

# n per Site × Trait (observations used in the full GxY model)
gxy_n <- purrr::map_dfr(SITES, function(site) {
  purrr::map_dfr(TRAITS, function(tr) {
    tibble(Site = site, Trait = tr,
           n = nobs(gxy_models[[site]][[tr]]$full))
  })
})

# Format each cell as "vcov (±SD, percent%)", with "~0.00%" for tiny values
fmt_cell <- function(vcov, sdcor, pct) {
  pct_str <- ifelse(pct < 0.01 & pct > 0, "~0.00%",
                    sprintf("%.2f%%", pct))
  sprintf("%.3f (±%.2f, %s)", vcov, sdcor, pct_str)
}

gxy_varcomp_table <- gxy_varcomp %>%
  mutate(cell = fmt_cell(vcov, sdcor, percent)) %>%
  dplyr::select(Site, Trait, grp, cell) %>%
  pivot_wider(names_from = grp, values_from = cell) %>%
  inner_join(gxy_n, by = c("Site", "Trait")) %>%
  mutate(`Population (n)` = paste0(Site, " (n = ", n, ")"),
         Trait = recode(Trait, !!!TRAIT_LABELS_SHORT),
         Trait = factor(Trait, levels = unname(TRAIT_LABELS_SHORT)),
         Site  = factor(Site,  levels = SITES)) %>%
  arrange(Site, Trait) %>%
  dplyr::select(`Population (n)`, Trait,
                `Recipient σ²`     = Recipient,
                `Donor σ²`         = Donor,
                `Donor:Year σ²`    = `Donor:Year`,
                `Residual σ²`      = Residual)

knitr::kable(gxy_varcomp_table, format = "html",
             align = c("l", "l", "r", "r", "r", "r"),
             caption = "Table 4. Variance components from the G×Y model (Eq. 4).") %>%
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "bordered", "condensed"),
    full_width = FALSE, position = "left"
  )
```

# 6. Multivariate plasticity (mean z-scored PIv across traits per family)

```{r mean-piv-family}
mean_PIv_by_family <- purrr::map_dfr(SITES, function(site) {
  plasticity[[site]]$PIv %>%
    filter(Trait %in% PIv_TRAITS) %>%
    group_by(Trait) %>%
    mutate(PIv_z = as.numeric(scale(PIv))) %>%   # z-score within site × trait
    ungroup() %>%
    group_by(Donor) %>%
    summarise(mean_PIv = mean(PIv_z, na.rm = TRUE), .groups = "drop") %>%
    mutate(Site = site, .before = 1)
})
mean_PIv_by_family
```

# 7. Paternal-family fitness W (two metrics)

We build two fitness data frames in parallel. Each has the columns
`Site`, `Donor`, `W_Y1`, `W_Y2`, `deltaW`, `mean_W`, so the same
analysis helper (defined in §8) runs on both.

## 7a. Primary: aster sire effects on the mean-value scale

```{r load-BV}
read_BV <- function(site, year) {
  fpath <- here("chapter_1_adaptive_capacity", "aster", "processed_data",
                paste0(site, year, "_bhat.Donor.mu.csv"))
  d <- read.csv(fpath, stringsAsFactors = FALSE, check.names = FALSE)
  colnames(d)[1] <- "raw"
  colnames(d)[2] <- "BV_W"
  d %>%
    as_tibble() %>%
    mutate(Donor = sub(".*Donor", "", raw),
           Site  = site, Year = paste0("20", year)) %>%
    dplyr::select(Site, Year, Donor, BV_W)
}

BV_W <- purrr::map_dfr(
  SITES, ~ bind_rows(read_BV(.x, "22"), read_BV(.x, "23")))

W_aster <- BV_W %>%
  pivot_wider(names_from = Year, values_from = BV_W,
              names_prefix = "W_") %>%
  rename(W_Y1 = W_2022, W_Y2 = W_2023) %>%
  mutate(deltaW = W_Y2 - W_Y1,
         mean_W = rowMeans(across(c(W_Y1, W_Y2)), na.rm = TRUE))
W_aster
```

## 7b. Sensitivity: observed family-mean fecundity

`est_fecundity = total_fruits × mean_seeds_per_fruit` per fitness
plant (Chapter 2 convention); we take the per-family mean each year.
Plants that died before flowering contribute zero (handled in §1).

```{r family-W-fec}
fec_family <- purrr::map_dfr(SITES, function(site) {
  site_data[[site]] %>%
    group_by(Donor, Year) %>%
    summarise(W    = mean(est_fecundity, na.rm = TRUE),
              W_n  = sum(!is.na(est_fecundity)),
              W_sd = sd(est_fecundity,   na.rm = TRUE),
              .groups = "drop") %>%
    mutate(W_se = W_sd / sqrt(pmax(W_n, 1)),
           Site = site, .before = 1)
})

W_fec <- fec_family %>%
  dplyr::select(Site, Donor, Year, W) %>%
  pivot_wider(names_from = Year, values_from = W,
              names_prefix = "W_") %>%
  rename(W_Y1 = W_2022, W_Y2 = W_2023) %>%
  mutate(deltaW = W_Y2 - W_Y1,
         mean_W = rowMeans(across(c(W_Y1, W_Y2)), na.rm = TRUE))
W_fec
```

# 8. Plasticity vs fitness — shared analysis helper

Both fitness data frames share the same column names (`W_Y1`, `W_Y2`,
`deltaW`, `mean_W`).  We define one helper that runs Eqs 5, 6, 7/8,
and the directional-plasticity-vs-mean-W Pearson correlation; it is
applied to `W_aster` (§9) and `W_fec` (§10).

```{r analysis-helper}
run_W_analyses <- function(W_df, plasticity, mean_dir, mean_piv) {

  # 8a. Pearson r: directional plasticity vs mean_W
  dir_meanW_cor <- purrr::map_dfr(SITES, function(site) {
    d <- plasticity[[site]]$directional %>%
      inner_join(filter(W_df, Site == site), by = "Donor")
    purrr::map_dfr(TRAITS, function(tr) {
      sub <- filter(d, Trait == tr)
      ct  <- with(sub, cor.test(directional, mean_W))
      tibble(Site = site, Trait = tr,
             n = sum(complete.cases(sub$directional, sub$mean_W)),
             r = unname(ct$estimate), p = ct$p.value)
    })
  })
  dir_meanW_multi <- purrr::map_dfr(SITES, function(site) {
    d <- filter(mean_dir, Site == site) %>%
      inner_join(filter(W_df, Site == site), by = "Donor")
    ct <- with(d, cor.test(mean_directional, mean_W))
    tibble(Site = site, Trait = "multivariate",
           n = sum(complete.cases(d$mean_directional, d$mean_W)),
           r = unname(ct$estimate), p = ct$p.value)
  })
  dir_meanW_cor <- bind_rows(dir_meanW_cor, dir_meanW_multi) %>%
    arrange(Site, Trait)

  # 8b. Eq. 5: W_Y2 ~ W_Y1 + directional
  eq5 <- purrr::map_dfr(SITES, function(site) {
    bw <- filter(W_df, Site == site)
    pl <- plasticity[[site]]$directional
    purrr::map_dfr(TRAITS, function(tr) {
      d <- pl %>% filter(Trait == tr) %>% inner_join(bw, by = "Donor")
      m <- lm(W_Y2 ~ W_Y1 + directional, data = d)
      tdy <- broom::tidy(m) %>% filter(term == "directional")
      tibble(Site = site, Trait = tr, n = nobs(m),
             beta_plast = tdy$estimate, p_plast = tdy$p.value,
             R2 = summary(m)$r.squared)
    })
  })
  eq5_multi <- purrr::map_dfr(SITES, function(site) {
    bw <- filter(W_df, Site == site)
    d  <- filter(mean_dir, Site == site) %>% inner_join(bw, by = "Donor")
    m  <- lm(W_Y2 ~ W_Y1 + mean_directional, data = d)
    tdy <- broom::tidy(m) %>% filter(term == "mean_directional")
    tibble(Site = site, Trait = "multivariate", n = nobs(m),
           beta_plast = tdy$estimate, p_plast = tdy$p.value,
           R2 = summary(m)$r.squared)
  })
  eq5 <- bind_rows(eq5, eq5_multi) %>% arrange(Site, Trait)

  # 8c. Eq. 6: deltaW ~ mean_W + directional
  eq6 <- purrr::map_dfr(SITES, function(site) {
    bw <- filter(W_df, Site == site)
    pl <- plasticity[[site]]$directional
    purrr::map_dfr(TRAITS, function(tr) {
      d <- pl %>% filter(Trait == tr) %>% inner_join(bw, by = "Donor")
      m <- lm(deltaW ~ mean_W + directional, data = d)
      tdy <- broom::tidy(m) %>% filter(term == "directional")
      tibble(Site = site, Trait = tr, n = nobs(m),
             beta_plast = tdy$estimate, p_plast = tdy$p.value,
             R2 = summary(m)$r.squared)
    })
  })
  eq6_multi <- purrr::map_dfr(SITES, function(site) {
    bw <- filter(W_df, Site == site)
    d  <- filter(mean_dir, Site == site) %>% inner_join(bw, by = "Donor")
    m  <- lm(deltaW ~ mean_W + mean_directional, data = d)
    tdy <- broom::tidy(m) %>% filter(term == "mean_directional")
    tibble(Site = site, Trait = "multivariate", n = nobs(m),
           beta_plast = tdy$estimate, p_plast = tdy$p.value,
           R2 = summary(m)$r.squared)
  })
  eq6 <- bind_rows(eq6, eq6_multi) %>% arrange(Site, Trait)

  # 8d. Eq. 7/8: deltaW ~ PIv (linear + quadratic check)
  run_piv_lm <- function(d, piv_col) {
    f_lin  <- as.formula(paste0("deltaW ~ ", piv_col))
    f_quad <- as.formula(paste0("deltaW ~ ", piv_col, " + I(", piv_col, "^2)"))
    m_lin  <- lm(f_lin,  data = d)
    m_quad <- lm(f_quad, data = d)
    q <- broom::tidy(m_quad) %>% filter(grepl("\\^2", term))
    l <- broom::tidy(m_lin)  %>% slice(2)
    tibble(n        = nobs(m_lin),
           p_quad   = q$p.value,
           beta_lin = l$estimate,
           p_lin    = l$p.value,
           R2_lin   = summary(m_lin)$r.squared,
           R2_quad  = summary(m_quad)$r.squared)
  }
  eq78 <- purrr::map_dfr(SITES, function(site) {
    bw  <- filter(W_df, Site == site)
    piv <- plasticity[[site]]$PIv
    purrr::map_dfr(PIv_TRAITS, function(tr) {
      d <- piv %>% filter(Trait == tr) %>% inner_join(bw, by = "Donor")
      run_piv_lm(d, "PIv") %>% mutate(Site = site, Trait = tr, .before = 1)
    })
  })
  eq78_multi <- purrr::map_dfr(SITES, function(site) {
    d <- inner_join(filter(W_df, Site == site),
                    filter(mean_piv, Site == site),
                    by = c("Site", "Donor"))
    run_piv_lm(d, "mean_PIv") %>%
      mutate(Site = site, Trait = "multivariate", .before = 1)
  })
  eq78 <- bind_rows(eq78, eq78_multi) %>% arrange(Site, Trait)

  list(dir_meanW_cor = dir_meanW_cor,
       eq5           = eq5,
       eq6           = eq6,
       eq78          = eq78)
}
```

# 9. PRIMARY analysis — aster sire effects

```{r run-primary}
primary <- run_W_analyses(W_aster, plasticity,
                          mean_directional_by_family,
                          mean_PIv_by_family)

cat("\n--- 9a. Pearson r: directional plasticity vs mean W (aster) ---\n")
print(primary$dir_meanW_cor, n = Inf)

cat("\n--- 9b. Eq. 5: W_Y2 ~ W_Y1 + directional (aster) ---\n")
print(primary$eq5, n = Inf)

cat("\n--- 9c. Eq. 6: deltaW ~ mean_W + directional (aster) ---\n")
print(primary$eq6, n = Inf)

cat("\n--- 9d. Eq. 7/8: deltaW ~ PIv (aster) ---\n")
print(primary$eq78, n = Inf)
```

# 10. SENSITIVITY analysis — observed fecundity

```{r run-sensitivity}
sensitivity <- run_W_analyses(W_fec, plasticity,
                              mean_directional_by_family,
                              mean_PIv_by_family)

cat("\n--- 10a. Pearson r: directional plasticity vs mean W (fecundity) ---\n")
print(sensitivity$dir_meanW_cor, n = Inf)

cat("\n--- 10b. Eq. 5 (fecundity) ---\n")
print(sensitivity$eq5, n = Inf)

cat("\n--- 10c. Eq. 6 (fecundity) ---\n")
print(sensitivity$eq6, n = Inf)

cat("\n--- 10d. Eq. 7/8 (fecundity) ---\n")
print(sensitivity$eq78, n = Inf)
```

# 11. Comparison: primary vs sensitivity

For each Site × Trait × test, we flag whether each metric (aster vs
fecundity) reached p < 0.05.  Cells where the two metrics agree on
significance reinforce a real effect; cells where only one is
significant indicate a result that should be treated as marginal under
the multiple-testing burden of 4 traits × 4 sites × multiple model
frameworks.

```{r comparison}
sig_compare <- function(prim, sens, value_col, p_col, label) {
  prim %>%
    dplyr::select(Site, Trait, primary_beta = !!sym(value_col),
                  primary_p = !!sym(p_col)) %>%
    inner_join(
      sens %>% dplyr::select(Site, Trait,
                             sens_beta = !!sym(value_col),
                             sens_p   = !!sym(p_col)),
      by = c("Site", "Trait")
    ) %>%
    mutate(test = label,
           sig_primary = primary_p < 0.05,
           sig_sens    = sens_p    < 0.05,
           agreement   = case_when(
             sig_primary &  sig_sens ~ "both p<0.05",
             sig_primary & !sig_sens ~ "primary only",
             !sig_primary & sig_sens ~ "sensitivity only",
             TRUE                    ~ "neither"
           )) %>%
    dplyr::select(test, Site, Trait, primary_beta, primary_p,
                  sens_beta, sens_p, agreement)
}

cmp_eq5  <- sig_compare(primary$eq5, sensitivity$eq5,
                        "beta_plast", "p_plast", "Eq. 5")
cmp_eq6  <- sig_compare(primary$eq6, sensitivity$eq6,
                        "beta_plast", "p_plast", "Eq. 6")
cmp_eq78 <- sig_compare(primary$eq78, sensitivity$eq78,
                        "beta_lin", "p_lin", "Eq. 8 (linear)")
cmp_cor  <- sig_compare(primary$dir_meanW_cor, sensitivity$dir_meanW_cor,
                        "r", "p", "Pearson r")

comparison <- bind_rows(cmp_cor, cmp_eq5, cmp_eq6, cmp_eq78) %>%
  arrange(test, Site, Trait)

cat("\n--- All Site × Trait × test comparisons ---\n")
print(comparison, n = Inf)

cat("\n--- Cells where the two metrics disagree on significance ---\n")
disagree <- comparison %>% filter(agreement %in% c("primary only",
                                                   "sensitivity only"))
print(disagree, n = Inf)

cat("\n--- Cells where both metrics reached p < 0.05 ---\n")
agree_sig <- comparison %>% filter(agreement == "both p<0.05")
print(agree_sig, n = Inf)
```

# 12. Figures

## Figure 5: Trait reaction norms per population

Each line is one paternal family; line colour distinguishes whether the
trait increased or decreased from Year 1 to Year 2.

```{r fig5, fig.width=9, fig.height=10}
reaction_norm_plot <- function(site) {
  fm <- plasticity[[site]]$fam_means
  dir <- fm %>%
    pivot_wider(names_from = Year, values_from = mean,
                names_prefix = "Y") %>%
    mutate(dir = if_else(Y2023 > Y2022, "increase", "decrease")) %>%
    dplyr::select(Donor, Trait, dir)

  plot_df <- fm %>%
    left_join(dir, by = c("Donor", "Trait")) %>%
    mutate(Trait_lab = recode(Trait, !!!TRAIT_LABELS))

  ggplot(plot_df,
         aes(Year, mean, group = Donor, colour = dir)) +
    geom_line(alpha = 0.5) + geom_point(size = 1) +
    facet_wrap(~ Trait_lab, scales = "free_y", ncol = 2) +
    scale_colour_manual(values = c(increase = "#1f78b4",
                                   decrease = "#e31a1c"),
                        name = NULL) +
    labs(title = paste0(site, " — paternal-family reaction norms"),
         x = NULL, y = "Family mean") +
    theme_bw() +
    theme(legend.position = "top",
          strip.background = element_blank())
}

fig5 <- wrap_plots(purrr::map(SITES, reaction_norm_plot), ncol = 2) +
  plot_annotation(tag_levels = "A")
fig5
```

## Figure 6: Multivariate plasticity trajectories (SLA × δ13C)

```{r fig6, fig.width=10, fig.height=8}
multivar_trajectory <- function(site) {
  fm <- plasticity[[site]]$fam_means %>%
    filter(Trait %in% c("SLA_SEG", "d13C_SEG")) %>%
    pivot_wider(names_from = Trait, values_from = mean)

  ggplot(fm, aes(SLA_SEG, d13C_SEG, group = Donor, colour = Year)) +
    geom_path(arrow = arrow(length = unit(2, "mm"), type = "closed"),
              alpha = 0.6) +
    geom_point(size = 1.4) +
    labs(title = site, x = "SLA (mm²/mg)", y = "δ13C (‰)") +
    theme_bw()
}

fig6 <- wrap_plots(purrr::map(SITES, multivar_trajectory), ncol = 2) +
  plot_annotation(tag_levels = "A")
fig6
```

## Figure 7: Variance partitioning (G, M, G×Y, residual)

```{r fig7, fig.width=10, fig.height=6}
varcomp_labels <- c(`Donor:Year` = "Donor × Year (G × Y)",
                    Donor        = "Donor (G)",
                    Recipient    = "Recipient (M)",
                    Residual     = "Residual")

fig7 <- gxy_varcomp %>%
  mutate(Component = recode(grp, !!!varcomp_labels),
         Component = factor(Component, levels = unname(varcomp_labels)),
         Trait_lab = recode(Trait, !!!TRAIT_LABELS),
         Site      = factor(Site, levels = SITES)) %>%
  ggplot(aes(Site, percent, fill = Component)) +
  geom_col(position = "stack", width = 0.7) +
  facet_wrap(~ Trait_lab, ncol = 2) +
  scale_fill_brewer(palette = "Set2") +
  labs(y = "Variance explained (%)", x = NULL) +
  theme_bw() + theme(legend.position = "bottom")
fig7
```

## Figure 8: Directional plasticity vs mean fitness (primary + sensitivity)

Top set of panels: primary analysis using aster sire effects.
Bottom set of panels: sensitivity analysis using observed fecundity.
Blue regression lines indicate p < 0.05 within that metric.

```{r fig8, fig.width=14, fig.height=18}
make_fig8_panel <- function(site, W_df, sig_df, ylab_text) {
  d_trait <- plasticity[[site]]$directional %>%
    inner_join(filter(W_df, Site == site), by = "Donor") %>%
    mutate(Trait_lab = recode(Trait, !!!TRAIT_LABELS),
           x = directional)

  d_multi <- filter(mean_directional_by_family, Site == site) %>%
    inner_join(filter(W_df, Site == site) %>%
                 dplyr::select(Donor, mean_W),
               by = "Donor") %>%
    mutate(Trait = "multivariate", Trait_lab = "Multivariate",
           x = mean_directional) %>%
    dplyr::select(Site, Donor, Trait, Trait_lab, x, mean_W)

  sig_keys <- sig_df %>% filter(p < 0.05) %>%
    mutate(key = paste(Site, Trait)) %>% pull(key)

  d <- bind_rows(
    d_trait %>% dplyr::select(Site, Donor, Trait, Trait_lab, x, mean_W),
    d_multi
  ) %>%
    mutate(Trait_lab = factor(Trait_lab,
                               levels = c("Multivariate",
                                          unname(TRAIT_LABELS))),
           sig = paste(site, Trait) %in% sig_keys)

  ggplot(d, aes(x, mean_W)) +
    geom_point(alpha = 0.7, size = 1.2) +
    geom_smooth(data = filter(d, sig),
                method = "lm", se = TRUE, colour = "steelblue") +
    facet_wrap(~ Trait_lab, scales = "free_x", ncol = 3) +
    labs(title = site,
         x = "Directional plasticity (Year 1 → Year 2)",
         y = ylab_text) +
    theme_bw() +
    theme(strip.background = element_blank())
}

fig8_primary <- wrap_plots(
  purrr::map(SITES, ~ make_fig8_panel(.x, W_aster, primary$dir_meanW_cor,
                                       expression(bar(W) ~ "(aster sire effect)"))),
  ncol = 2) +
  plot_annotation(title = "Primary analysis — aster sire effects",
                  tag_levels = "A")

fig8_sens <- wrap_plots(
  purrr::map(SITES, ~ make_fig8_panel(.x, W_fec, sensitivity$dir_meanW_cor,
                                       expression(bar(W) ~ "(observed fecundity)"))),
  ncol = 2) +
  plot_annotation(title = "Sensitivity analysis — observed fecundity",
                  tag_levels = "A")

fig8_primary
fig8_sens
```

# Session info

```{r}
sessionInfo()
```
