---
title: "Chapter_3_GxE_backup"
author: "Helen Payne"
date: "2026-01-09"
output: html_document
---

\-\-- title: "Ch3_GxE" author: "Helen Payne" date: "2025-12-03" output: html_document editor_options: markdown: wrap: 72 \-\--

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}
library(tidyverse)
```

# ANGELO COAST

Read in the data

```{r}
AC_22 <- read_csv(here::here("data_sheets/compiled_sheets/AC_compiledsheet_full_2022.csv"))
AC_23 <- read_csv(here::here("data_sheets/compiled_sheets/AC_compiledsheet_full_2023.csv"))
```

Create a dataset

```{r}
ac22_trait <- AC_22 %>%
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2022") #adding year column

ac23_trait <- AC_23 %>%
  filter(Gen == "G1") %>%    # just include generation 1
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2023") #adding year column

# combine years
ac_traits <- bind_rows(ac22_trait, ac23_trait) %>%
  filter(F_plant == "TRUE")
```

Get paternal family means

```{r}
fam_means <- ac_traits %>%
  group_by(Donor, Year) %>% #all calculations done separately for each paternal family within each year
  summarise(
    across(
      c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
      ~ mean(.x, na.rm = TRUE), # get the family mean corolla diameter for each Donor × Year, remove NAs
      .names = "mean_{.col}"
    ),
    .groups = "drop" #ungroups the results
  )
```

Assign colors for Increasing and Decreasing

```{r}
dirs_wide <- fam_means %>%
  pivot_wider(
    names_from = Year,
    values_from = starts_with("mean_"),
    names_sep = "_"
  ) %>%
  mutate(
    dir_corolla_diam_mm_SEG = if_else(mean_corolla_diam_mm_SEG_2023 >
                                        mean_corolla_diam_mm_SEG_2022,
                                      "increase", "decrease"),
    dir_SLA_SEG  = if_else(mean_SLA_SEG_2023  > mean_SLA_SEG_2022,
                           "increase", "decrease"),
    dir_d13C_SEG = if_else(mean_d13C_SEG_2023 > mean_d13C_SEG_2022,
                           "increase", "decrease"),
    dir_msm_all  = if_else(mean_msm_all_2023  > mean_msm_all_2022,
                           "increase", "decrease")
  ) %>%
  select(Donor, starts_with("dir_"))

#Long format for plotting
fam_long <- fam_means %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to  = "Trait",
    values_to = "mean_value"
  )

dirs_long <- dirs_wide %>%
  pivot_longer(
    cols = starts_with("dir_"),
    names_to  = "Trait",
    values_to = "dir"
  ) %>%
  mutate(
    Trait = case_when(
      Trait == "dir_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "dir_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "dir_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "dir_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  )
```

Plot the data

```{r}
plot_dat <- fam_long %>%
  mutate(
    Trait = case_when(
      Trait == "mean_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "mean_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "mean_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "mean_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  ) %>%
  left_join(dirs_long, by = c("Donor", "Trait"))

#One plot, all traits, blue = increase, red = decrease
p <- ggplot(plot_dat,
            aes(x = Year, y = mean_value,
                group = Donor, color = dir)) +
  geom_line(alpha = 0.9) +
  geom_point(size = 2.5) +
  scale_x_discrete(expand = c(0.05, 0)) +
  scale_color_manual(
    values = c(increase = "#2c7fb8", decrease = "#d95f0e"),
    name   = "Change 2022 → 2023"
  ) +
  facet_wrap(~ Trait, scales = "free_y") +
  labs(x = "Year", y = "AC paternal family mean") +
  theme_bw(base_size = 18) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 16),
    axis.text.y  = element_text(size = 16)
  )

p

ggsave("Figures/AC_reaction_norms_traits.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Pivot the dataframe so each family shows trait means

```{r}
fam_wide <- fam_means %>%
  tidyr::pivot_wider(
    id_cols = Donor,
    names_from = Year,
    values_from = c(mean_corolla_diam_mm_SEG, mean_SLA_SEG, mean_d13C_SEG, mean_msm_all)
  )
```

## **1. Is there plasticity in the traits of interest (corolla diameter, sla, d13C, msm)?**

YES. Spearman's Rank Correlations. We will also observe plasticity using the plasticity index as well.

```{r}
# Spearman correlation for corolla area
cor.test(fam_wide$mean_corolla_diam_mm_SEG_2022, fam_wide$mean_corolla_diam_mm_SEG_2023, method = "spearman")
# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_corolla_diam_mm_SEG_2022 and fam_wide$mean_corolla_diam_mm_SEG_2023
# S = 8882, p-value = 0.3025
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.1667917 


# Spearman correlation for SLA
cor.test(fam_wide$mean_SLA_SEG_2022, fam_wide$mean_SLA_SEG_2023, method = "spearman")
# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_SLA_SEG_2022 and fam_wide$mean_SLA_SEG_2023
# S = 6828, p-value = 0.2574
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.1906117 

# Spearman correlation for d13C
cor.test(fam_wide$mean_d13C_SEG_2022, fam_wide$mean_d13C_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_d13C_SEG_2022 and fam_wide$mean_d13C_SEG_2023
# S = 7792, p-value = 0.1959
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#      rho 
# 0.211336

# Spearman correlation for msm
cor.test(fam_wide$mean_msm_all_2022, fam_wide$mean_msm_all_2023, method = "spearman")

# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_msm_all_2022 and fam_wide$mean_msm_all_2023
# S = 7790, p-value = 0.1954
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.2115385
```

## **2. Do traits differ in how plastic they are, on average, across genotypes?**

Calculate how much two traits differ from one another using PIv (Valladares et al., 2006):

```{r}
piv_df <- fam_long %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_mean = max(mean_value, na.rm = TRUE), #determine the maximum donor mean
    min_mean = min(mean_value, na.rm = TRUE), #determine the minimum donor mean
    PIv = (max_mean - min_mean) / max_mean, #calculate the PIv to determine how much traits differ from eachother
    .groups = "drop"
  )
```

Calculate plasticity by trait using PIv:

```{r}
mean_piv_by_trait <- piv_df %>%
  group_by(Trait) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))

# mean_SLA_SEG
# 0.12840040

# mean_corolla_diam_mm_SEG
# 0.08422690

# mean_d13C_SEG
# -0.01515385

# mean_msm_all
# 0.15994703
```

Calculate how much traits differ from one another using Plmd (used for non-normal distributions):

SLA and msm were slightly non-normal based on histograms earlier.

```{r}

ac_traits_long <- ac_traits %>%
  pivot_longer(
    cols = c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
    names_to = "Trait",
    values_to = "value"
  )

PImd_df <- ac_traits_long %>%
  group_by(Donor, Trait, Year) %>%
  summarise(
    median_value = median(value, na.rm = TRUE),
    .groups = "drop"
  )

PImd_df <- PImd_df %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_median = max(median_value, na.rm = TRUE),
    min_median = min(median_value, na.rm = TRUE),
    PImd = (max_median - min_median) / max_median,
    .groups = "drop"
  )
```

Do traits differ in how plastic they are, on average, across genotypes? Calculate plasticity by trait:

```{r}
mean_PImd_by_trait <- PImd_df %>%
  group_by(Trait) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))


#SLA_SEG
# 0.10921634
# 
# corolla_diam_mm_SEG
# 0.09407846
# 
# d13C_SEG
# -0.01753768
# 
# msm_all
# 0.15127971
```

Comparing PIv and PImd:

```{r}

cor.test(mean_PImd_by_trait$mean_PImd, mean_piv_by_trait$mean_PIv)
#Families ranked as highly plastic using PIv are also ranked as highly plastic using PImd

# 	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_trait$mean_PImd and mean_piv_by_trait$mean_PIv
# t = 9.1336, df = 2, p-value = 0.01178
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.5402505 0.9997650
# sample estimates:
#       cor 
# 0.9882242

#Traits ranked as highly plastic using PIv are also ranked as highly plastic using PImd

#PIv and PImd are esentially the same
```

## **3. Within each trait, do paternal families differ in their plastic responses (i.e., is there genetic variation for plasticity)?**

Do paternal families differ in their plastic responses to corolla area?

```{r}
library(lme4)
library(lmerTest)

model <- lmer(corolla_diam_mm_SEG ~ Year * Donor + (1|Recipient), data = ac_traits)
anova(model)

# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value   Pr(>F)   
# Year        62.82  62.818     1 329.56 10.4689 0.001337 **
# Donor      388.58   9.964    39  76.16  1.6605 0.029776 * 
# Year:Donor 320.56   8.220    39 319.11  1.3698 0.076932 . 
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
# 	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99558, p-value = 0.2707
plot(model)
hist(ac_traits$corolla_diam_mm_SEG) #normal
```

Do paternal families differ in their plastic responses to SLA?

```{r}

model <- lmer(SLA_SEG ~ Year * Donor + (1|Recipient), data = ac_traits)
anova(model)
# 
# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value   Pr(>F)   
# Year       205.43 205.431     1 323.98  8.1552 0.004571 **
# Donor      812.25  20.827    39  61.95  0.8268 0.734852   
# Year:Donor 814.30  22.619    36 314.31  0.8979 0.640354

shapiro.test(residuals(model)) #not normal
# 	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.79037, p-value < 2.2e-16
plot(model)
hist(ac_traits$SLA_SEG) #not normal
```

Do paternal families differ in their plastic responses to d13C?

```{r}

model <- lmer(d13C_SEG ~ Year * Donor + (1|Recipient), data = ac_traits)
anova(model)

# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value    Pr(>F)    
# Year       11.999 11.9993     1 362.92 25.9319 5.694e-07 ***
# Donor      37.458  0.9605    39  57.87  2.0757  0.005692 ** 
# Year:Donor 27.296  0.7183    38 350.33  1.5524  0.023198 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99783, p-value = 0.8157
plot(model)
hist(ac_traits$d13C_SEG) #normal
```

Do paternal families differ in their plastic responses to msm?

```{r}

model <- lmer(msm_all ~ Year * Donor + (1|Recipient), data = ac_traits) #warning because there are a few missing datapoints, but its ok
anova(model)

#Type III Analysis of Variance Table with Satterthwaite's method
#             Sum Sq Mean Sq NumDF  DenDF F value Pr(>F)
# Year        0.0408 0.04081     1 338.31  0.1018 0.7498
# Donor      18.9839 0.48677    39  65.97  1.2145 0.2399
# Year:Donor 16.4686 0.43338    38 334.84  1.0813 0.3483

shapiro.test(residuals(model))

plot(model)
hist(ac_traits$msm_all) #normal
```

## **4. Across traits, are some paternal families consistently more plastic than others?**

Calculate the overall multivariate plasticity using PIv:

```{r}
mean_piv_by_family <- piv_df %>%
  group_by(Donor) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))
```

Calculate the overall multivariate plasticity using PImd:

```{r}
mean_PImd_by_family <- PImd_df %>%
  group_by(Donor) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))
```

Compare and Contrast PImd and PIv:

```{r}
cor.test(mean_PImd_by_family$mean_PImd, mean_piv_by_family$mean_PIv)

# 	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_family$mean_PImd and mean_piv_by_family$mean_PIv
# t = 6.655, df = 38, p-value = 7.231e-08
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.5471672 0.8507222
# sample estimates:
#       cor 
# 0.7336309 

#Families ranked as highly plastic using PIv are also ranked as highly plastic using PImd
```

## **5. How is a paternal family's trait-specific plasticity as well as it's cross-trait (average) plasticity associated with breeding values for fitness?**

```{r}

AC22_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/AC22_bhat.Donor.mu.csv")
AC22_BV <- AC22_BV %>%
  rename(Donor = X) %>% 
  rename(BV_W = bhat.Donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))


AC23_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/AC23_bhat.Donor.mu.csv")
AC23_BV <- AC23_BV %>%
  rename(Donor = X) %>%   
  rename(BV_W = bhat.donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))

```

Merged the two datasets by Donor for 2022

```{r}
merged <- AC22_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated for 2022

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = -0.90835, df = 38, p-value = 0.3694
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.4374244  0.1736142
# sample estimates:
#        cor 
# -0.1457792 
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity for 2022

```{r}
library(dplyr)
library(ggplot2)

# --- Merge datasets by Donor ---
merged <- AC22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
   mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

# --- Scatterplot ---
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall Plasticity",
    y = "AC BV(W) 2022"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/AC_BV(W)2022_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Merged the two datasets by Donor for 2023

```{r}
merged <- AC23_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated in 2023

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = -1.6439, df = 38, p-value = 0.1085
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.52688526  0.05854156
# sample estimates:
#        cor 
# -0.2576667
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity in 2023

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- AC23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

# --- Scatterplot ---
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall plasticity",
    y = "AC BV(W) 2023"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/AC_BV(W)2023_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

#### Determine if breeding values are correlated with any specific traits:

See if BV(W) in 2022 and Plasticity in corolla area are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- AC22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.41572, df = 38, p-value = 0.68
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2494516  0.3710190
# sample estimates:
#        cor 
# 0.06728636
```

See if BV(W) and Plasticity in SLA are are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- AC22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)


# 	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = 0.61824, df = 38, p-value = 0.5401
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2185109  0.3989002
# sample estimates:
#        cor 
# 0.09979103
```

See if BV(W) and Plasticity in SLA are are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- AC22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)


# 	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -0.47999, df = 38, p-value = 0.634
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3799502  0.2396763
# sample estimates:
#        cor 
# -0.0776289 

```

See if BV(W) and Plasticity in SLA are are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- AC22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)


# 	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -1.6989, df = 38, p-value = 0.0975
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.53308632  0.04994371
# sample estimates:
#        cor 
# -0.2656994
```

Determine if breeding values are correlated with any specific traits: See if BV(W) in 2023 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- AC23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = -0.49489, df = 38, p-value = 0.6235
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3820105  0.2374034
# sample estimates:
#       cor 
# -0.080024
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- AC23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = 0.49014, df = 38, p-value = 0.6269
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2381278  0.3813546
# sample estimates:
#        cor 
# 0.07926108 
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- AC23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -0.9775, df = 38, p-value = 0.3345
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.4463482  0.1628389
# sample estimates:
#        cor 
# -0.1566149 
```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- AC23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)


# 	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -1.9321, df = 38, p-value = 0.06082
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.55855720  0.01369765
# sample estimates:
#        cor 
# -0.2990879
```

Linear Regression to determine if a correlation exists:

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and absolute plasticity are correlated:

```{r}

model_linear  <- lm(BV_W ~ mean_PIv, data = merged)
model_quad    <- lm(BV_W ~ mean_PIv + I(mean_PIv^2), data = merged)

anova(model_linear, model_quad)

#Analysis of Variance Table
# 
# Model 1: BV_W ~ mean_PIv
# Model 2: BV_W ~ mean_PIv + I(mean_PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     38 13.077                           
# 2     37 12.718  1   0.35936 1.0455 0.3132

##the linear model is better than the quadratic

summary(model_linear)
#Call:
# lm(formula = BV_W ~ mean_PIv, data = merged)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -1.22727 -0.39430 -0.06287  0.25538  1.57531 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   6.2505     0.2475  25.255   <2e-16 ***
# mean_PIv     -4.2213     2.5679  -1.644    0.108    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.5866 on 38 degrees of freedom
# Multiple R-squared:  0.06639,	Adjusted R-squared:  0.04182 
# F-statistic: 2.702 on 1 and 38 DF,  p-value: 0.1085

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear...

shapiro.test(residuals(model_linear))
shapiro.test(residuals(model_quad))

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in corolla diameter are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_corolla)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_corolla)

anova(model_linear, model_quad)
# Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     38 13.918                           
# 2     37 13.816  1   0.10162 0.2721  0.605

summary(model_linear)

#Call:
# lm(formula = BV_W ~ PIv, data = merged_corolla)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.93325 -0.48914 -0.01214  0.31044  1.68933 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   5.9383     0.1624  36.556   <2e-16 ***
# PIv          -0.7713     1.5585  -0.495    0.624    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.6052 on 38 degrees of freedom
# Multiple R-squared:  0.006404,	Adjusted R-squared:  -0.01974 
# F-statistic: 0.2449 on 1 and 38 DF,  p-value: 0.6235

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear..

shapiro.test(residuals(model_linear))
shapiro.test(residuals(model_quad))

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in d13C are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_d13C)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_d13C)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     38 13.664                           
# 2     37 13.122  1   0.54194 1.5281 0.2242

summary(model_linear)

#Call:
# lm(formula = BV_W ~ PIv, data = merged_d13C)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.88707 -0.46341 -0.02474  0.32059  1.63525 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   5.7678     0.1436  40.160   <2e-16 ***
# PIv          -6.9587     7.1188  -0.978    0.335    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.5996 on 38 degrees of freedom
# Multiple R-squared:  0.02453,	Adjusted R-squared:  -0.001142 
# F-statistic: 0.9555 on 1 and 38 DF,  p-value: 0.3345

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear))
shapiro.test(residuals(model_quad))

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in msm are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_msm)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_msm)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     38 18.379                           
# 2     37 17.976  1   0.40332 0.8302 0.3681

summary(model_linear)

#Call:
# lm(formula = BV_W ~ PIv, data = merged_msm)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -1.0401 -0.4851 -0.1099  0.3792  2.4943 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  15.1371     0.1782  84.927   <2e-16 ***
# PIv          -1.4900     0.8770  -1.699   0.0975 .  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.6955 on 38 degrees of freedom
# Multiple R-squared:  0.0706,	Adjusted R-squared:  0.04614 
# F-statistic: 2.886 on 1 and 38 DF,  p-value: 0.0975

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad)

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in SLA are correlated:

```{r}
library(car)
library(lmtest)


model_linear  <- lm(BV_W ~ PIv, data = merged_SLA)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_SLA)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq     F Pr(>F)
# 1     38 13.919                          
# 2     37 13.769  1   0.15034 0.404 0.5289

summary(model_linear)

#
# Call:
# lm(formula = BV_W ~ PIv, data = merged_SLA)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.86524 -0.48217 -0.02293  0.29135  1.70735 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   5.8076     0.1646   35.28   <2e-16 ***
# PIv           0.5113     1.0432    0.49    0.627    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.6052 on 38 degrees of freedom
# Multiple R-squared:  0.006282,	Adjusted R-squared:  -0.01987 
# F-statistic: 0.2402 on 1 and 38 DF,  p-value: 0.6269

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear))
shapiro.test(residuals(model_quad))

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```
# BODEGA BAY

Read in the data

```{r}
BB_22 <- read_csv(here::here("data_sheets/compiled_sheets/BB_compiledsheet_full_2022.csv"))
BB_23 <- read_csv(here::here("data_sheets/compiled_sheets/BB_compiledsheet_full_2023.csv"))
```

Create a dataset

```{r}
BB22_trait <- BB_22 %>%
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2022") #adding year column

BB23_trait <- BB_23 %>%
  filter(Gen == "G1") %>%    # just include generation 1
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2023") #adding year column

# combine years
BB_traits <- bind_rows(BB22_trait, BB23_trait) %>%
  filter(F_plant == "TRUE")
```

Get paternal family means

```{r}
fam_means <- BB_traits %>%
  group_by(Donor, Year) %>% #all calculations done separately for each paternal family within each year
  summarise(
    across(
      c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
      ~ mean(.x, na.rm = TRUE), # get the family mean corolla diameter for each Donor × Year, remove NAs
      .names = "mean_{.col}"
    ),
    .groups = "drop" #ungroups the results
  )

```

Assign colors for Increasing and Decreasing

```{r}
dirs_wide <- fam_means %>%
  pivot_wider(
    names_from = Year,
    values_from = starts_with("mean_"),
    names_sep = "_"
  ) %>%
  mutate(
    dir_corolla_diam_mm_SEG = if_else(mean_corolla_diam_mm_SEG_2023 >
                                        mean_corolla_diam_mm_SEG_2022,
                                      "increase", "decrease"),
    dir_SLA_SEG  = if_else(mean_SLA_SEG_2023  > mean_SLA_SEG_2022,
                           "increase", "decrease"),
    dir_d13C_SEG = if_else(mean_d13C_SEG_2023 > mean_d13C_SEG_2022,
                           "increase", "decrease"),
    dir_msm_all  = if_else(mean_msm_all_2023  > mean_msm_all_2022,
                           "increase", "decrease")
  ) %>%
  select(Donor, starts_with("dir_"))

#Long format for plotting
fam_long <- fam_means %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to  = "Trait",
    values_to = "mean_value"
  )

dirs_long <- dirs_wide %>%
  pivot_longer(
    cols = starts_with("dir_"),
    names_to  = "Trait",
    values_to = "dir"
  ) %>%
  mutate(
    Trait = case_when(
      Trait == "dir_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "dir_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "dir_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "dir_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  )
```

Plot the data

```{r}
plot_dat <- fam_long %>%
  mutate(
    Trait = case_when(
      Trait == "mean_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "mean_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "mean_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "mean_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  ) %>%
  left_join(dirs_long, by = c("Donor", "Trait"))

#One plot, all traits, blue = increase, red = decrease
p <- ggplot(plot_dat,
            aes(x = Year, y = mean_value,
                group = Donor, color = dir)) +
  geom_line(alpha = 0.9) +
  geom_point(size = 2.5) +
  scale_x_discrete(expand = c(0.05, 0)) +
  scale_color_manual(
    values = c(increase = "#2c7fb8", decrease = "#d95f0e"),
    name   = "Change 2022 → 2023"
  ) +
  facet_wrap(~ Trait, scales = "free_y") +
  labs(x = "Year", y = "BB paternal family mean") +
  theme_bw(base_size = 18) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 16),
    axis.text.y  = element_text(size = 16)
  )

p

ggsave("Figures/BB_reaction_norms_traits.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Pivot the dataframe so each family shows trait means
```{r}
fam_wide <- fam_means %>%
  tidyr::pivot_wider(
    id_cols = Donor,
    names_from = Year,
    values_from = c(mean_corolla_diam_mm_SEG, mean_SLA_SEG, mean_d13C_SEG, mean_msm_all)
  )
```

## **1. Is there plasticity in the traits of interest (corolla diameter, SLA, d13C, msm)?**

YES. Spearman's Rank Correlations. We will also observe plasticity using the plasticity index as well.

```{r}
# Spearman correlation for corolla area
cor.test(fam_wide$mean_corolla_diam_mm_SEG_2022, fam_wide$mean_corolla_diam_mm_SEG_2023, method = "spearman")
# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_corolla_diam_mm_SEG_2022 and fam_wide$mean_corolla_diam_mm_SEG_2023
# S = 11644, p-value = 0.004065
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.4059184


# Spearman correlation for SLA **SIGNIFICANT CORRELATION**
cor.test(fam_wide$mean_SLA_SEG_2022, fam_wide$mean_SLA_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_SLA_SEG_2022 and fam_wide$mean_SLA_SEG_2023
# S = 22138, p-value = 0.169
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# -0.2015849

# Spearman correlation for d13C
cor.test(fam_wide$mean_d13C_SEG_2022, fam_wide$mean_d13C_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_d13C_SEG_2022 and fam_wide$mean_d13C_SEG_2023
# S = 12931, p-value = 0.01675
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.3402638 

# Spearman correlation for msm
cor.test(fam_wide$mean_msm_all_2022, fam_wide$mean_msm_all_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_msm_all_2022 and fam_wide$mean_msm_all_2023
# S = 16228, p-value = 0.6792
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# 0.06174838 
```

## **2. Do traits differ in how plastic they are, on average, across genotypes?**

Calculate how much two traits differ from one another using PIv (Valladares et al., 2006):

```{r}
piv_df <- fam_long %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_mean = max(mean_value, na.rm = TRUE), #determine the maximum donor mean
    min_mean = min(mean_value, na.rm = TRUE), #determine the minimum donor mean
    PIv = (max_mean - min_mean) / max_mean, #calculate the PIv to determine how much traits differ from eachother
    .groups = "drop"
  )
```

Calculate plasticity by trait using PIv:

```{r}
mean_piv_by_trait <- piv_df %>%
  group_by(Trait) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))

# mean_SLA_SEG
# 0.19641035
# 2
# mean_corolla_diam_mm_SEG
# 0.10031997
# 3
# mean_d13C_SEG
# -0.02074321
# 4
# mean_msm_all
# 0.26916701


#Mean seed mass is most plastic
```

Calculate how much traits differ from one another using Plmd (used for non-normal distributions):

SLA and msm were slightly non-normal based on histograms earlier.

```{r}

BB_traits_long <- BB_traits %>%
  pivot_longer(
    cols = c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
    names_to = "Trait",
    values_to = "value"
  )

PImd_df <- BB_traits_long %>%
  group_by(Donor, Trait, Year) %>%
  summarise(
    median_value = median(value, na.rm = TRUE),
    .groups = "drop"
  )

PImd_df <- PImd_df %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_median = max(median_value, na.rm = TRUE),
    min_median = min(median_value, na.rm = TRUE),
    PImd = (max_median - min_median) / max_median,
    .groups = "drop"
  )
```

Do traits differ in how plastic they are, on average, across genotypes? Calculate plasticity by trait:

```{r}
mean_PImd_by_trait <- PImd_df %>%
  group_by(Trait) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))


#SLA_SEG
# 0.18273880
# 
# corolla_diam_mm_SEG
# 0.11388103
# 
# d13C_SEG
# -0.02491601
# 
# msm_all
# 0.27613576
```

Comparing PIv and PImd:
```{r}
cor.test(mean_PImd_by_trait$mean_PImd, mean_piv_by_trait$mean_PIv)
#Families ranked as highly plastic using PIv are also ranked as highly plastic using PImd

#Pearson's product-moment correlation
# 
# data:  mean_PImd_by_trait$mean_PImd and mean_piv_by_trait$mean_PIv
# t = 14.847, df = 2, p-value = 0.004506
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.7956510 0.9999104
# sample estimates:
#       cor 
# 0.9954939

#Traits ranked as highly plastic using PIv are also ranked as highly plastic using PImd

#PIv and PImd are esentially the same
```

## **3. Within each trait, do paternal families differ in their plastic responses (i.e., is there genetic variation for plasticity)?**

Do paternal families differ in their plastic responses to corolla area?

```{r}
library(lme4)
library(lmerTest)

model <- lmer(corolla_diam_mm_SEG ~ Year * Donor + (1|Recipient), data = BB_traits)
anova(model) #Singular


#Type III Analysis of Variance Table with Satterthwaite's method
#             Sum Sq Mean Sq NumDF DenDF F value    Pr(>F)    
# Year        673.46  673.46     1   505 33.0771 1.536e-08 ***
# Donor      1251.45   26.07    48   505  1.2805    0.1048    
# Year:Donor  608.97   12.69    48   505  0.6231    0.9780    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99463, p-value = 0.03256 #not good
plot(model)
hist(BB_traits$corolla_diam_mm_SEG) #normal
```

Do paternal families differ in their plastic responses to SLA?

```{r}

model <- lmer(SLA_SEG ~ Year * Donor + (1|Recipient), data = BB_traits)
anova(model)
#Type III Analysis of Variance Table with Satterthwaite's method
#             Sum Sq Mean Sq NumDF   DenDF F value  Pr(>F)  
# Year         15.63  15.632     1 215.978  0.4574 0.49956  
# Donor      1649.24  34.359    48  75.521  1.0053 0.48414  
# Year:Donor 2180.03  46.384    47 200.923  1.3572 0.07803 .
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model)) #not normal
#
# 	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.95724, p-value = 2.596e-08
plot(model)
hist(BB_traits$SLA_SEG) #normal-ish
```

Do paternal families differ in their plastic responses to d13C?

```{r}

model <- lmer(d13C_SEG ~ Year * Donor + (1|Recipient), data = BB_traits)
anova(model)

# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value    Pr(>F)    
# Year       12.869 12.8689     1 366.82 11.6504 0.0007135 ***
# Donor      69.344  1.4447    48  81.03  1.3079 0.1422165    
# Year:Donor 47.880  0.9975    48 348.31  0.9031 0.6577596    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99519, p-value = 0.132
#normal
plot(model)
hist(BB_traits$d13C_SEG) #normal
```

Do paternal families differ in their plastic responses to msm?

```{r}

model <- lmer(msm_all ~ Year * Donor + (1|Recipient), data = BB_traits) #warning because there are a few missing datapoints, but its ok
anova(model)

#Type III Analysis of Variance Table with Satterthwaite's method
#             Sum Sq Mean Sq NumDF   DenDF F value    Pr(>F)    
# Year        16.316 16.3162     1 220.758  6.8521 0.0094661 ** 
# Donor      226.207  4.7126    48  84.712  1.9791 0.0030229 ** 
# Year:Donor 214.904  4.6718    46 210.842  1.9620 0.0007473 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.95733, p-value = 1.602e-08
#not normal
plot(model)
hist(BB_traits$msm_all) #normal-ish?
```

## **4. Across traits, are some paternal families consistently more plastic than others?**

Calculate the overall multivariate plasticity using PIv:

```{r}
mean_piv_by_family <- piv_df %>%
  group_by(Donor) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))
```

Calculate the overall multivariate plasticity using PImd:

```{r}
mean_PImd_by_family <- PImd_df %>%
  group_by(Donor) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))
```

```{r}
cor.test(mean_PImd_by_family$mean_PImd, mean_piv_by_family$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_family$mean_PImd and mean_piv_by_family$mean_PIv
# t = 8.0419, df = 47, p-value = 2.205e-10
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.6104362 0.8584909
# sample estimates:
#       cor 
# 0.7610044 

```

## **5. How is a paternal family's trait-specific plasticity as well as it's cross-trait (average) plasticity associated with breeding values for fitness?**

```{r}

BB22_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/BB22_bhat.Donor.mu.csv")
BB22_BV <- BB22_BV %>%
  rename(Donor = X) %>% 
  rename(BV_W = bhat.Donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))


BB23_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/BB23_bhat.Donor.mu.csv")
BB23_BV <- BB23_BV %>%
  rename(Donor = X) %>%   
  rename(BV_W = bhat.donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))

```

Merged the two datasets by Donor for 2022

```{r}
merged <- BB22_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated for 2022

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = -0.9125, df = 47, p-value = 0.3662
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3983559  0.1550090
# sample estimates:
#        cor 
# -0.1319386
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity for 2022

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- BB22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
   mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

#plot
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall Plasticity",
    y = "BB BV(W) 2022"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/BB_BV(W)2022_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Merged the two datasets by Donor for 2023

```{r}
merged <- BB23_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated in 2023

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = -0.41562, df = 47, p-value = 0.6796
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3359920  0.2245041
# sample estimates:
#         cor 
# -0.06051267 
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity in 2023

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- BB23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

# --- Scatterplot ---
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall plasticity",
    y = "BB BV(W) 2023"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/BB_BV(W)2023_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

#### Determine if breeding values are correlated with any specific traits:

See if BV(W) in 2022 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- BB22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.85579, df = 47, p-value = 0.3965
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1630055  0.3914313
# sample estimates:
#       cor 
# 0.1238688
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- BB22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = 0.015751, df = 47, p-value = 0.9875
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2790794  0.2833110
# sample estimates:
#         cor 
# 0.002297453
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- BB22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = 0.77348, df = 47, p-value = 0.4431
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1745888  0.3812881
# sample estimates:
#       cor 
# 0.1121118 

```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- BB22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)


#Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -1.4127, df = 47, p-value = 0.1643
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.45708126  0.08414693
# sample estimates:
#        cor 
# -0.2018252
```

Determine if breeding values are correlated with any specific traits: See if BV(W) in 2023 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- BB23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation

# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.6797, df = 47, p-value = 0.5
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1877447  0.3696032
# sample estimates:
#        cor 
# 0.09866051
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- BB23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = -0.69204, df = 47, p-value = 0.4923
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3711482  0.1860166
# sample estimates:
#        cor 
# -0.1004334 
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- BB23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -0.98754, df = 47, p-value = 0.3284
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.4074373  0.1444107
# sample estimates:
#        cor 
# -0.1425756
```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- BB23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -0.13999, df = 47, p-value = 0.8893
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2998908  0.2622865
# sample estimates:
#         cor 
# -0.02041582 
```

Linear Regression to determine if a correlation exists: \####################################################### Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and absolute plasticity are correlated:

```{r}

model_linear  <- lm(BV_W ~ mean_PIv, data = merged)
model_quad    <- lm(BV_W ~ mean_PIv + I(mean_PIv^2), data = merged)

anova(model_linear, model_quad)

#Analysis of Variance Table
# 
# Model 1: BV_W ~ mean_PIv
# Model 2: BV_W ~ mean_PIv + I(mean_PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     47 564.34                           
# 2     46 561.02  1    3.3269 0.2728  0.604

summary(model_linear)
#Call:
# lm(formula = BV_W ~ mean_PIv, data = merged)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -4.5257 -2.5273 -0.6813  1.9686 11.4419 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   15.902      1.324  12.013 6.23e-16 ***
# mean_PIv      -3.744      9.008  -0.416     0.68    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.465 on 47 degrees of freedom
# Multiple R-squared:  0.003662,	Adjusted R-squared:  -0.01754 
# F-statistic: 0.1727 on 1 and 47 DF,  p-value: 0.6796

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear...

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in corolla diameter are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_corolla)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_corolla)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df   RSS Df  Sum of Sq  F Pr(>F)
# 1     47 560.9                        
# 2     46 560.9  1 0.00028351  0 0.9962

summary(model_linear)

# Call:
# lm(formula = BV_W ~ PIv, data = merged_corolla)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -4.6103 -2.7410 -0.5098  2.4246 11.0299 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  14.8606     0.9247   16.07   <2e-16 ***
# PIv           5.2986     7.7954    0.68      0.5    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.455 on 47 degrees of freedom
# Multiple R-squared:  0.009734,	Adjusted R-squared:  -0.01134 
# F-statistic: 0.462 on 1 and 47 DF,  p-value: 0.5

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear..

shapiro.test(residuals(model_linear)) #non-normal
shapiro.test(residuals(model_quad)) #non-normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in d13C are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_d13C)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_d13C)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F  Pr(>F)  
# 1     47 554.90                              
# 2     46 493.38  1    61.526 5.7364 0.02075 *
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

summary(model_linear)
#Call:
# lm(formula = BV_W ~ PIv, data = merged_d13C)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -5.0145 -2.2300 -0.7544  2.6733  9.9128 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  14.6849     0.8683  16.913   <2e-16 ***
# PIv         -34.0959    34.5262  -0.988    0.328    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.436 on 47 degrees of freedom
# Multiple R-squared:  0.02033,	Adjusted R-squared:  -0.0005163 
# F-statistic: 0.9752 on 1 and 47 DF,  p-value: 0.3284

summary(model_quad)

# Call:
# lm(formula = BV_W ~ PIv + I(PIv^2), data = merged_d13C)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -6.5289 -2.3982 -0.1377  2.0149  8.2976 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   16.249      1.054  15.413   <2e-16 ***
# PIv          160.436     87.635   1.831   0.0736 .  
# I(PIv^2)    3907.155   1631.324   2.395   0.0207 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.275 on 46 degrees of freedom
# Multiple R-squared:  0.129,	Adjusted R-squared:  0.09108 
# F-statistic: 3.405 on 2 and 46 DF,  p-value: 0.04178

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)


#THE QUADRATIC MODEL DETECTS OPPTOMISING SELECTION FOR d13C
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in msm are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_msm)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_msm)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F  Pr(>F)  
# 1     47 566.18                              
# 2     46 519.09  1    47.091 4.1731 0.04683 *
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

summary(model_linear)

#
# Call:
# lm(formula = BV_W ~ PIv, data = merged_msm)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -4.6059 -2.7105 -0.7424  2.0012 11.3889 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  15.4894     0.8538   18.14   <2e-16 ***
# PIv          -0.3615     2.5825   -0.14    0.889    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.471 on 47 degrees of freedom
# Multiple R-squared:  0.0004168,	Adjusted R-squared:  -0.02085 
# F-statistic: 0.0196 on 1 and 47 DF,  p-value: 0.8893


summary(model_quad)

#Call:
# lm(formula = BV_W ~ PIv + I(PIv^2), data = merged_msm)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -5.2188 -1.9986 -0.7142  1.9959 10.5248 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   14.067      1.081  13.017   <2e-16 ***
# PIv           13.959      7.443   1.876   0.0671 .  
# I(PIv^2)     -22.250     10.892  -2.043   0.0468 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.359 on 46 degrees of freedom
# Multiple R-squared:  0.08356,	Adjusted R-squared:  0.04371 
# F-statistic: 2.097 on 2 and 46 DF,  p-value: 0.1344

#THE QUADRATIC MODEL DETECTS OPPTOMISING SELECTION FOR msm

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in SLA are correlated:

```{r}
library(car)
library(lmtest)


model_linear  <- lm(BV_W ~ PIv, data = merged_SLA)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_SLA)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     47 560.70                           
# 2     46 558.87  1    1.8357 0.1511 0.6993

summary(model_linear)
#Call:
# lm(formula = BV_W ~ PIv, data = merged_SLA)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -4.9663 -2.5562 -0.5188  2.2206 11.3274 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  15.8874     0.8693  18.277   <2e-16 ***
# PIv          -2.5215     3.6436  -0.692    0.492    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 3.454 on 47 degrees of freedom
# Multiple R-squared:  0.01009,	Adjusted R-squared:  -0.01098 
# F-statistic: 0.4789 on 1 and 47 DF,  p-value: 0.4923

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```


# BLUE OAK

Read in the data

```{r}
BO_22 <- read_csv(here::here("data_sheets/compiled_sheets/BO_compiledsheet_full_2022.csv"))
BO_23 <- read_csv(here::here("data_sheets/compiled_sheets/BO_compiledsheet_full_2023.csv"))
```

Create a dataset

```{r}
BO22_trait <- BO_22 %>%
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2022") #adding year column

BO23_trait <- BO_23 %>%
  filter(Gen == "G1") %>%    # just include generation 1
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2023") #adding year column

#REMOVING THIS ONE ODD OUTLIER WHICH WAS LIKELY A MISTAKE
BO23_trait <- BO23_trait %>%
  mutate(SLA_SEG = ifelse(Donor == "BO_194" &
                          Recipient == "BO_098",
                          NA,
                          SLA_SEG))

# combine years
BO_traits <- bind_rows(BO22_trait, BO23_trait) %>%
  filter(F_plant == "TRUE")
```

Get paternal family means

```{r}
fam_means <- BO_traits %>%
  group_by(Donor, Year) %>% #all calculations done separately for each paternal family within each year
  summarise(
    across(
      c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
      ~ mean(.x, na.rm = TRUE), # get the family mean corolla diameter for each Donor × Year, remove NAs
      .names = "mean_{.col}"
    ),
    .groups = "drop" #ungroups the results
  )

```

Assign colors for Increasing and Decreasing

```{r}
dirs_wide <- fam_means %>%
  pivot_wider(
    names_from = Year,
    values_from = starts_with("mean_"),
    names_sep = "_"
  ) %>%
  mutate(
    dir_corolla_diam_mm_SEG = if_else(mean_corolla_diam_mm_SEG_2023 >
                                        mean_corolla_diam_mm_SEG_2022,
                                      "increase", "decrease"),
    dir_SLA_SEG  = if_else(mean_SLA_SEG_2023  > mean_SLA_SEG_2022,
                           "increase", "decrease"),
    dir_d13C_SEG = if_else(mean_d13C_SEG_2023 > mean_d13C_SEG_2022,
                           "increase", "decrease"),
    dir_msm_all  = if_else(mean_msm_all_2023  > mean_msm_all_2022,
                           "increase", "decrease")
  ) %>%
  select(Donor, starts_with("dir_"))

#Long format for plotting
fam_long <- fam_means %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to  = "Trait",
    values_to = "mean_value"
  )

dirs_long <- dirs_wide %>%
  pivot_longer(
    cols = starts_with("dir_"),
    names_to  = "Trait",
    values_to = "dir"
  ) %>%
  mutate(
    Trait = case_when(
      Trait == "dir_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "dir_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "dir_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "dir_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  )
```

Plot the data

```{r}
plot_dat <- fam_long %>%
  mutate(
    Trait = case_when(
      Trait == "mean_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "mean_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "mean_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "mean_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  ) %>%
  left_join(dirs_long, by = c("Donor", "Trait"))

#One plot, all traits, blue = increase, red = decrease
p <- ggplot(plot_dat,
            aes(x = Year, y = mean_value,
                group = Donor, color = dir)) +
  geom_line(alpha = 0.9) +
  geom_point(size = 2.5) +
  scale_x_discrete(expand = c(0.05, 0)) +
  scale_color_manual(
    values = c(increase = "#2c7fb8", decrease = "#d95f0e"),
    name   = "Change 2022 → 2023"
  ) +
  facet_wrap(~ Trait, scales = "free_y") +
  labs(x = "Year", y = "BO paternal family mean") +
  theme_bw(base_size = 18) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 16),
    axis.text.y  = element_text(size = 16)
  )

p

ggsave("Figures/BO_reaction_norms_traits.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Pivot the dataframe so each family shows trait means

```{r}
fam_wide <- fam_means %>%
  tidyr::pivot_wider(
    id_cols = Donor,
    names_from = Year,
    values_from = c(mean_corolla_diam_mm_SEG, mean_SLA_SEG, mean_d13C_SEG, mean_msm_all)
  )
```

## **1. Is there plasticity in the traits of interest (corolla diameter, SLA, d13C, msm)?**

YES. Spearman's Rank Correlations. We will also observe plasticity using the plasticity index as well.

```{r}
# Spearman correlation for corolla area
cor.test(fam_wide$mean_corolla_diam_mm_SEG_2022, fam_wide$mean_corolla_diam_mm_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_corolla_diam_mm_SEG_2022 and fam_wide$mean_corolla_diam_mm_SEG_2023
# S = 10578, p-value = 0.3655
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.1428571 


# Spearman correlation for SLA **SIGNIFICANT CORRELATION**
cor.test(fam_wide$mean_SLA_SEG_2022, fam_wide$mean_SLA_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_SLA_SEG_2022 and fam_wide$mean_SLA_SEG_2023
# S = 8736, p-value = 0.02604
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.3403805

# Spearman correlation for d13C
cor.test(fam_wide$mean_d13C_SEG_2022, fam_wide$mean_d13C_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_d13C_SEG_2022 and fam_wide$mean_d13C_SEG_2023
# S = 12052, p-value = 0.3289
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.1506854

# Spearman correlation for msm
cor.test(fam_wide$mean_msm_all_2022, fam_wide$mean_msm_all_2023, method = "spearman")
#	Spearman's rank correlation rho

# data:  fam_wide$mean_msm_all_2022 and fam_wide$mean_msm_all_2023
# S = 10372, p-value = 0.547
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#        rho 
# 0.09651568 
```

## **2. Do traits differ in how plastic they are, on average, across genotypes?**

Calculate how much two traits differ from one another using PIv (Valladares et al., 2006):

```{r}
piv_df <- fam_long %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_mean = max(mean_value, na.rm = TRUE), #determine the maximum donor mean
    min_mean = min(mean_value, na.rm = TRUE), #determine the minimum donor mean
    PIv = (max_mean - min_mean) / max_mean, #calculate the PIv to determine how much traits differ from eachother
    .groups = "drop"
  )
```

Calculate plasticity by trait using PIv:

```{r}
mean_piv_by_trait <- piv_df %>%
  group_by(Trait) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))

# mean_SLA_SEG
# 0.11008189
# 
# mean_corolla_diam_mm_SEG
# 0.11343378
# 
# mean_d13C_SEG
# -0.04560326
# 
# mean_msm_all
# 0.28343208

#Mean seed mass is most plastic
```

Calculate how much traits differ from one another using Plmd (used for non-normal distributions):

SLA and msm were slightly non-normal based on histograms earlier.

```{r}

BO_traits_long <- BO_traits %>%
  pivot_longer(
    cols = c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
    names_to = "Trait",
    values_to = "value"
  )

PImd_df <- BO_traits_long %>%
  group_by(Donor, Trait, Year) %>%
  summarise(
    median_value = median(value, na.rm = TRUE),
    .groups = "drop"
  )

PImd_df <- PImd_df %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_median = max(median_value, na.rm = TRUE),
    min_median = min(median_value, na.rm = TRUE),
    PImd = (max_median - min_median) / max_median,
    .groups = "drop"
  )
```

Do traits differ in how plastic they are, on average, across genotypes? Calculate plasticity by trait:

```{r}
mean_PImd_by_trait <- PImd_df %>%
  group_by(Trait) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))

# 
# SLA_SEG
# 0.10258560
# 
# corolla_diam_mm_SEG
# 0.11685655
# 
# d13C_SEG
# -0.04541372
# 
# msm_all
# 0.28806691
```

Comparing PIv and PImd:

```{r}
cor.test(mean_PImd_by_trait$mean_PImd, mean_piv_by_trait$mean_PIv)
#Families ranked as highly plastic using PIv are also ranked as highly plastic using PImd

# 	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_trait$mean_PImd and mean_piv_by_trait$mean_PIv
# t = 37.809, df = 2, p-value = 0.0006988
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.9653800 0.9999861
# sample estimates:
#       cor 
# 0.9993012

#Traits ranked as highly plastic using PIv are also ranked as highly plastic using PImd

#PIv and PImd are esentially the same
```

## **3. Within each trait, do paternal families differ in their plastic responses (i.e., is there genetic variation for plasticity)?**

Do paternal families differ in their plastic responses to corolla area?

```{r}
library(lme4)
library(lmerTest)

model <- lmer(corolla_diam_mm_SEG ~ Year * Donor + (1|Recipient), data = BO_traits)
anova(model)

# Missing cells for: Year2023:DonorBO_032, Year2023:DonorBO_151.  
# Interpret type III hypotheses with care.
# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF   DenDF F value  Pr(>F)  
# Year        74.93  74.928     1 286.677  5.3354 0.02161 *
# Donor      769.56  17.897    43  84.787  1.2744 0.17054  
# Year:Donor 754.82  18.410    41 264.015  1.3109 0.10870  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
# 	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.9966, p-value = 0.5695
plot(model)
hist(BO_traits$corolla_diam_mm_SEG) #normal
```

Do paternal families differ in their plastic responses to SLA?

```{r}

model <- lmer(SLA_SEG ~ Year * Donor + (1|Recipient), data = BO_traits)
anova(model)
#Missing cells for: Year2023:DonorBO_106.  
# Interpret type III hypotheses with care.
# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF   DenDF F value Pr(>F)
# Year         8.77  8.7749     1 200.516  0.4272 0.5141
# Donor      773.59 17.9905    43  50.084  0.8759 0.6700
# Year:Donor 423.75 10.0894    42 182.587  0.4912 0.9961

shapiro.test(residuals(model)) #not normal
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.83646, p-value < 2.2e-16
plot(model)
hist(BO_traits$SLA_SEG) #not normal
```

Do paternal families differ in their plastic responses to d13C?

```{r}

model <- lmer(d13C_SEG ~ Year * Donor + (1|Recipient), data = BO_traits)
anova(model)

# #Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF   DenDF F value   Pr(>F)    
# Year       73.629  73.629     1 225.022 72.4513 2.44e-15 ***
# Donor      86.492   2.011    43  65.242  1.9793 0.006216 ** 
# Year:Donor 72.961   1.697    43 203.880  1.6696 0.010055 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.98972, p-value = 0.02196
#not normal
plot(model)
hist(BO_traits$d13C_SEG) #not normal
```

Do paternal families differ in their plastic responses to msm?

```{r}

model <- lmer(msm_all ~ Year * Donor + (1|Recipient), data = BO_traits) #warning because there are a few missing datapoints, but its ok
anova(model)

#Type III Analysis of Variance Table with Satterthwaite's method
#             Sum Sq Mean Sq NumDF DenDF F value  Pr(>F)  
# Year         0.301  0.3013     1   155  0.1101 0.74044  
# Donor      168.108  3.9095    43   155  1.4291 0.06006 .
# Year:Donor 134.419  3.3605    40   155  1.2284 0.18850  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.96399, p-value = 9.597e-06
#not normal
plot(model)
hist(BO_traits$msm_all) #not normal
```

## **4. Across traits, are some paternal families consistently more plastic than others?**

Calculate the overall multivariate plasticity using PIv:

```{r}
mean_piv_by_family <- piv_df %>%
  group_by(Donor) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))
```

Calculate the overall multivariate plasticity using PImd:

```{r}
mean_PImd_by_family <- PImd_df %>%
  group_by(Donor) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))
```

```{r}
cor.test(mean_PImd_by_family$mean_PImd, mean_piv_by_family$mean_PIv)

# 	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_family$mean_PImd and mean_piv_by_family$mean_PIv
# t = 13.404, df = 42, p-value < 2.2e-16
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.8235307 0.9446824
# sample estimates:
#       cor 
# 0.9002973

```

## **5. How is a paternal family's trait-specific plasticity as well as it's cross-trait (average) plasticity associated with breeding values for fitness?**

```{r}

BO22_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/BO22_bhat.Donor.mu.csv")
BO22_BV <- BO22_BV %>%
  rename(Donor = X) %>% 
  rename(BV_W = bhat.Donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))


BO23_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/BO23_bhat.Donor.mu.csv")
BO23_BV <- BO23_BV %>%
  rename(Donor = X) %>%   
  rename(BV_W = bhat.donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))

```

Merged the two datasets by Donor for 2022

```{r}
merged <- BO22_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated for 2022

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = 0.50149, df = 42, p-value = 0.6186
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2248808  0.3656556
# sample estimates:
#        cor 
# 0.07715036  
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity for 2022

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- BO22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
   mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

#plot
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall Plasticity",
    y = "BO BV(W) 2022"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/BO_BV(W)2022_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Merged the two datasets by Donor for 2023

```{r}
merged <- BO23_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated in 2023

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = 0.77061, df = 42, p-value = 0.4453
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1853004  0.4009026
# sample estimates:
#       cor 
# 0.1180755 
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity in 2023

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- BO23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

# --- Scatterplot ---
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall plasticity",
    y = "BO BV(W) 2023"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/BO_BV(W)2023_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

#### Determine if breeding values are correlated with any specific traits:

See if BV(W) in 2022 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- BO22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.067739, df = 42, p-value = 0.9463
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2873201  0.3063815
# sample estimates:
#       cor 
# 0.0104518
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- BO22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = 0.40102, df = 42, p-value = 0.6904
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2395116  0.3521833
# sample estimates:
#        cor 
# 0.06176035 
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- BO22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = 0.60317, df = 42, p-value = 0.5496
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2099874  0.3791201
# sample estimates:
#        cor 
# 0.09267009 

```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- BO22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = 0.30336, df = 42, p-value = 0.7631
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2536436  0.3389332
# sample estimates:
#        cor 
# 0.04675767 
```

Determine if breeding values are correlated with any specific traits: See if BV(W) in 2023 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- BO23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.20309, df = 42, p-value = 0.84
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2680513  0.3251782
# sample estimates:
#        cor 
# 0.03132163 
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- BO23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = -0.1286, df = 42, p-value = 0.8983
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3148650  0.2786826
# sample estimates:
#         cor 
# -0.01983919 
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- BO23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -0.13753, df = 42, p-value = 0.8913
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3161057  0.2774113
# sample estimates:
#         cor 
# -0.02121644
```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- BO23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = 0.9425, df = 42, p-value = 0.3513
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1597900  0.4227361
# sample estimates:
#       cor 
# 0.1439177
```

Linear Regression to determine if a correlation exists: \####################################################### Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and absolute plasticity are correlated:

```{r}

model_linear  <- lm(BV_W ~ mean_PIv, data = merged)
model_quad    <- lm(BV_W ~ mean_PIv + I(mean_PIv^2), data = merged)

anova(model_linear, model_quad)

#Analysis of Variance Table
# 
# Model 1: BV_W ~ mean_PIv
# Model 2: BV_W ~ mean_PIv + I(mean_PIv^2)
#   Res.Df    RSS Df Sum of Sq     F Pr(>F)
# 1     42 2.0793                          
# 2     41 2.0793  1 9.017e-06 2e-04 0.9894

##the linear model is better than the quadratic

summary(model_linear)
#Call:
# lm(formula = BV_W ~ mean_PIv, data = merged)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.31353 -0.14221 -0.02386  0.08100  0.74839 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   1.7520     0.0754  23.235   <2e-16 ***
# mean_PIv      0.2936     0.5855   0.501    0.619    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.2225 on 42 degrees of freedom
# Multiple R-squared:  0.005952,	Adjusted R-squared:  -0.01772 
# F-statistic: 0.2515 on 1 and 42 DF,  p-value: 0.6186

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear...

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in corolla diameter are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_corolla)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_corolla)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df     RSS Df Sum of Sq      F Pr(>F)
# 1     42 0.51208                           
# 2     41 0.49935  1  0.012724 1.0448 0.3127

summary(model_linear)

# Call:
# lm(formula = BV_W ~ PIv, data = merged_corolla)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.14548 -0.08550 -0.02175  0.05684  0.34568 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  0.62990    0.03177  19.829   <2e-16 ***
# PIv          0.04844    0.23851   0.203     0.84    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.1104 on 42 degrees of freedom
# Multiple R-squared:  0.000981,	Adjusted R-squared:  -0.02281 
# F-statistic: 0.04124 on 1 and 42 DF,  p-value: 0.84

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear..

shapiro.test(residuals(model_linear)) #non-normal
shapiro.test(residuals(model_quad)) #non-normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in d13C are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_d13C)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_d13C)

anova(model_linear, model_quad)
#Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df     RSS Df Sum of Sq      F Pr(>F)
# 1     42 0.51235                           
# 2     41 0.50572  1 0.0066292 0.5374 0.4677

summary(model_linear)
# Call:
# lm(formula = BV_W ~ PIv, data = merged_d13C)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.14156 -0.08498 -0.02323  0.05564  0.34206 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  0.63190    0.03036  20.812   <2e-16 ***
# PIv         -0.07657    0.55675  -0.138    0.891    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.1104 on 42 degrees of freedom
# Multiple R-squared:  0.0004501,	Adjusted R-squared:  -0.02335 
# F-statistic: 0.01891 on 1 and 42 DF,  p-value: 0.8913

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in msm are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_msm)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_msm)

anova(model_linear, model_quad)
# Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df     RSS Df Sum of Sq      F  Pr(>F)  
# 1     42 0.50196                              
# 2     41 0.44756  1  0.054402 4.9837 0.03111 *
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

summary(model_linear)

#Call:
# lm(formula = BV_W ~ PIv, data = merged_msm)
# 
# Residuals:
#       Min        1Q    Median        3Q       Max 
# -0.135492 -0.087382 -0.007078  0.052905  0.307246 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  0.61212    0.02969  20.615   <2e-16 ***
# PIv          0.08213    0.08714   0.943    0.351    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.1093 on 42 degrees of freedom
# Multiple R-squared:  0.02071,	Adjusted R-squared:  -0.002604 
# F-statistic: 0.8883 on 1 and 42 DF,  p-value: 0.3513

summary(model_quad)

# Call:
# lm(formula = BV_W ~ PIv + I(PIv^2), data = merged_msm)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.18027 -0.07502 -0.03331  0.05905  0.20553 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  0.67230    0.03914  17.177   <2e-16 ***
# PIv         -0.44845    0.25184  -1.781   0.0824 .  
# I(PIv^2)     0.77695    0.34803   2.232   0.0311 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.1045 on 41 degrees of freedom
# Multiple R-squared:  0.1268,	Adjusted R-squared:  0.08425 
# F-statistic: 2.978 on 2 and 41 DF,  p-value: 0.062

#THE QUADRATIC MODEL DETECTS OPPTOMISING SELECTION FOR msm

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in SLA are correlated:

```{r}
library(car)
library(lmtest)


model_linear  <- lm(BV_W ~ PIv, data = merged_SLA)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_SLA)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df     RSS Df Sum of Sq      F Pr(>F)
# 1     42 0.51238                           
# 2     41 0.51096  1 0.0014157 0.1136 0.7378

summary(model_linear)
# Call:
# lm(formula = BV_W ~ PIv, data = merged_SLA)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.14456 -0.08564 -0.02018  0.05418  0.34543 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)  0.63813    0.02697  23.657   <2e-16 ***
# PIv         -0.02479    0.19277  -0.129    0.898    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.1105 on 42 degrees of freedom
# Multiple R-squared:  0.0003936,	Adjusted R-squared:  -0.02341 
# F-statistic: 0.01654 on 1 and 42 DF,  p-value: 0.8983

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

# HASTINGS RANCH

Read in the data

```{r}
HR_22 <- read_csv(here::here("data_sheets/compiled_sheets/HR_compiledsheet_full_2022.csv"))
HR_23 <- read_csv(here::here("data_sheets/compiled_sheets/HR_compiledsheet_full_2023.csv"))
```

Create a dataset

```{r}
HR22_trait <- HR_22 %>%
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2022") #adding year column

HR23_trait <- HR_23 %>%
  filter(Gen == "G1") %>%    # just include generation 1
  select(Donor, Recipient, F_plant, corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all) %>% #including maternal and paternal parents
  mutate(Year = "2023") #adding year column

#REMOVING THIS ONE ODD OUTLIER WHICH WAS LIKELY A MISTAKE
HR23_trait <- HR23_trait %>%
  mutate(SLA_SEG = ifelse(Donor == "HR_194" &
                          Recipient == "HR_098",
                          NA,
                          SLA_SEG))

# combine years
HR_traits <- bind_rows(HR22_trait, HR23_trait) %>%
  filter(F_plant == "TRUE")
```

Get paternal family means

```{r}
fam_means <- HR_traits %>%
  group_by(Donor, Year) %>% #all calculations done separately for each paternal family within each year
  summarise(
    across(
      c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
      ~ mean(.x, na.rm = TRUE), # get the family mean corolla diameter for each Donor × Year, remove NAs
      .names = "mean_{.col}"
    ),
    .groups = "drop" #ungroups the results
  )

```

Assign colors for Increasing and Decreasing

```{r}
dirs_wide <- fam_means %>%
  pivot_wider(
    names_from = Year,
    values_from = starts_with("mean_"),
    names_sep = "_"
  ) %>%
  mutate(
    dir_corolla_diam_mm_SEG = if_else(mean_corolla_diam_mm_SEG_2023 >
                                        mean_corolla_diam_mm_SEG_2022,
                                      "increase", "decrease"),
    dir_SLA_SEG  = if_else(mean_SLA_SEG_2023  > mean_SLA_SEG_2022,
                           "increase", "decrease"),
    dir_d13C_SEG = if_else(mean_d13C_SEG_2023 > mean_d13C_SEG_2022,
                           "increase", "decrease"),
    dir_msm_all  = if_else(mean_msm_all_2023  > mean_msm_all_2022,
                           "increase", "decrease")
  ) %>%
  select(Donor, starts_with("dir_"))

#Long format for plotting
fam_long <- fam_means %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to  = "Trait",
    values_to = "mean_value"
  )

dirs_long <- dirs_wide %>%
  pivot_longer(
    cols = starts_with("dir_"),
    names_to  = "Trait",
    values_to = "dir"
  ) %>%
  mutate(
    Trait = case_when(
      Trait == "dir_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "dir_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "dir_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "dir_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  )
```

Plot the data

```{r}
plot_dat <- fam_long %>%
  mutate(
    Trait = case_when(
      Trait == "mean_corolla_diam_mm_SEG" ~ "corolla diameter (mm)",
      Trait == "mean_SLA_SEG"             ~ "specific leaf area (mm/mg)",
      Trait == "mean_d13C_SEG"            ~ "δ13C (‰)",
      Trait == "mean_msm_all"             ~ "seed mass (mg)",
      TRUE ~ Trait
    )
  ) %>%
  left_join(dirs_long, by = c("Donor", "Trait"))

#One plot, all traits, blue = increase, red = decrease
p <- ggplot(plot_dat,
            aes(x = Year, y = mean_value,
                group = Donor, color = dir)) +
  geom_line(alpha = 0.9) +
  geom_point(size = 2.5) +
  scale_x_discrete(expand = c(0.05, 0)) +
  scale_color_manual(
    values = c(increase = "#2c7fb8", decrease = "#d95f0e"),
    name   = "Change 2022 → 2023"
  ) +
  facet_wrap(~ Trait, scales = "free_y") +
  labs(x = "Year", y = "HR paternal family mean") +
  theme_bw(base_size = 18) +
  theme(
    legend.position = "bottom",
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 16),
    axis.text.y  = element_text(size = 16)
  )

p

ggsave("Figures/HR_reaction_norms_traits.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Pivot the dataframe so each family shows trait means

```{r}
fam_wide <- fam_means %>%
  tidyr::pivot_wider(
    id_cols = Donor,
    names_from = Year,
    values_from = c(mean_corolla_diam_mm_SEG, mean_SLA_SEG, mean_d13C_SEG, mean_msm_all)
  )
```

## **1. Is there plasticity in the traits of interest (corolla diameter, SLA, d13C, msm)?**

YES. Spearman's Rank Correlations. We will also observe plasticity using the plasticity index as well.

```{r}
# Spearman correlation for corolla area
cor.test(fam_wide$mean_corolla_diam_mm_SEG_2022, fam_wide$mean_corolla_diam_mm_SEG_2023, method = "spearman")
#	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_corolla_diam_mm_SEG_2022 and fam_wide$mean_corolla_diam_mm_SEG_2023
# S = 13074, p-value = 0.1965
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.1937095 


# Spearman correlation for SLA **SIGNIFICANT CORRELATION**
cor.test(fam_wide$mean_SLA_SEG_2022, fam_wide$mean_SLA_SEG_2023, method = "spearman")

# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_SLA_SEG_2022 and fam_wide$mean_SLA_SEG_2023
# S = 15466, p-value = 0.9022
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#         rho 
# -0.01884058

# Spearman correlation for d13C
cor.test(fam_wide$mean_d13C_SEG_2022, fam_wide$mean_d13C_SEG_2023, method = "spearman")
# 	Spearman's rank correlation rho
# 
# data:  fam_wide$mean_d13C_SEG_2022 and fam_wide$mean_d13C_SEG_2023
# S = 5570.4, p-value = 3.065e-06
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.6330457 

# Spearman correlation for msm
cor.test(fam_wide$mean_msm_all_2022, fam_wide$mean_msm_all_2023, method = "spearman")

#	Spearman's rank correlation rho
#
# data:  fam_wide$mean_msm_all_2022 and fam_wide$mean_msm_all_2023
# S = 11998, p-value = 0.08106
# alternative hypothesis: true rho is not equal to 0
# sample estimates:
#       rho 
# 0.2600678
```

## **2. Do traits differ in how plastic they are, on average, across genotypes?**

Calculate how much two traits differ from one another using PIv (Valladares et al., 2006):

```{r}
piv_df <- fam_long %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_mean = max(mean_value, na.rm = TRUE), #determine the maximum donor mean
    min_mean = min(mean_value, na.rm = TRUE), #determine the minimum donor mean
    PIv = (max_mean - min_mean) / max_mean, #calculate the PIv to determine how much traits differ from eachother
    .groups = "drop"
  )
```

Calculate plasticity by trait using PIv:

```{r}
mean_piv_by_trait <- piv_df %>%
  group_by(Trait) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))


# mean_SLA_SEG
# 0.12559425
# 
# mean_corolla_diam_mm_SEG
# 0.05634288
# 
# mean_d13C_SEG
# -0.01180732
# 
# mean_msm_all
# 0.21584238


#Mean seed mass is most plastic
```

Calculate how much traits differ from one another using Plmd (used for non-normal distributions):

SLA and msm were slightly non-normal based on histograms earlier.

```{r}

HR_traits_long <- HR_traits %>%
  pivot_longer(
    cols = c(corolla_diam_mm_SEG, SLA_SEG, d13C_SEG, msm_all),
    names_to = "Trait",
    values_to = "value"
  )

PImd_df <- HR_traits_long %>%
  group_by(Donor, Trait, Year) %>%
  summarise(
    median_value = median(value, na.rm = TRUE),
    .groups = "drop"
  )

PImd_df <- PImd_df %>%
  group_by(Donor, Trait) %>%
  summarise(
    max_median = max(median_value, na.rm = TRUE),
    min_median = min(median_value, na.rm = TRUE),
    PImd = (max_median - min_median) / max_median,
    .groups = "drop"
  )
```

Do traits differ in how plastic they are, on average, across genotypes? Calculate plasticity by trait:

```{r}
mean_PImd_by_trait <- PImd_df %>%
  group_by(Trait) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))

# SLA_SEG
# 0.13051724
# 
# corolla_diam_mm_SEG
# 0.05848524
# 
# d13C_SEG
# -0.01212758
# 
# msm_all
# 0.21967071
```

Comparing PIv and PImd:

```{r}
cor.test(mean_PImd_by_trait$mean_PImd, mean_piv_by_trait$mean_PIv)
#Families ranked as highly plastic using PIv are also ranked as highly plastic using PImd

# 	Pearson's product-moment correlation
# 
# data:  mean_PImd_by_trait$mean_PImd and mean_piv_by_trait$mean_PIv
# t = 108.27, df = 2, p-value = 8.53e-05
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.9957101 0.9999983
# sample estimates:
#       cor 
# 0.9999147 

#Traits ranked as highly plastic using PIv are also ranked as highly plastic using PImd

#PIv and PImd are esentially the same
```

## **3. Within each trait, do paternal families differ in their plastic responses (i.e., is there genetic variation for plasticity)?**

Do paternal families differ in their plastic responses to corolla area?

```{r}
library(lme4)
library(lmerTest)

model <- lmer(corolla_diam_mm_SEG ~ Year * Donor + (1|Recipient), data = HR_traits)
anova(model) #Singular

#Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF DenDF F value    Pr(>F)    
# Year       110.06 110.060     1   575 13.9949 0.0002017 ***
# Donor      573.22  12.738    45   575  1.6197 0.0077630 ** 
# Year:Donor 322.19   7.160    45   575  0.9104 0.6405154    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99199, p-value = 0.001158

plot(model)
hist(HR_traits$corolla_diam_mm_SEG) #normal
```

Do paternal families differ in their plastic responses to SLA?

```{r}

model <- lmer(SLA_SEG ~ Year * Donor + (1|Recipient), data = HR_traits)
anova(model)
# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value   Pr(>F)   
# Year       295.04 295.044     1 433.31  9.9653 0.001707 **
# Donor      565.09  12.558    45  81.40  0.4241 0.998894   
# Year:Donor 588.46  13.374    44 415.94  0.4517 0.999111   
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model)) #not normal
#Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.70759, p-value < 2.2e-16
plot(model)
hist(HR_traits$SLA_SEG) #not normal
```

Do paternal families differ in their plastic responses to d13C?

```{r}

model <- lmer(d13C_SEG ~ Year * Donor + (1|Recipient), data = HR_traits)
anova(model)

#Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value    Pr(>F)    
# Year       10.501 10.5011     1 500.33 25.4254 6.440e-07 ***
# Donor      51.483  1.1441    45  92.08  2.7700 1.811e-05 ***
# Year:Donor 17.465  0.3969    44 494.87  0.9611    0.5458    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.99353, p-value = 0.007075
#not normal
plot(model)
hist(HR_traits$d13C_SEG) #not normal
```

Do paternal families differ in their plastic responses to msm?

```{r}

model <- lmer(msm_all ~ Year * Donor + (1|Recipient), data = HR_traits) #warning because there are a few missing datapoints, but its ok
anova(model)

# Type III Analysis of Variance Table with Satterthwaite's method
#            Sum Sq Mean Sq NumDF  DenDF F value    Pr(>F)    
# Year       21.758 21.7584     1 447.16 24.2113 1.216e-06 ***
# Donor      52.373  1.1638    45  93.49  1.2950   0.14736    
# Year:Donor 62.880  1.3973    45 440.37  1.5549   0.01486 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

shapiro.test(residuals(model))
#
# 	Shapiro-Wilk normality test
# 
# data:  residuals(model)
# W = 0.94147, p-value = 1.532e-14
#not normal
plot(model)
hist(HR_traits$msm_all) #not normal
```

## **4. Across traits, are some paternal families consistently more plastic than others?**

Calculate the overall multivariate plasticity using PIv:

```{r}
mean_piv_by_family <- piv_df %>%
  group_by(Donor) %>%
  summarise(mean_PIv = mean(PIv, na.rm = TRUE))
```

Calculate the overall multivariate plasticity using PImd:

```{r}
mean_PImd_by_family <- PImd_df %>%
  group_by(Donor) %>%
  summarise(mean_PImd = mean(PImd, na.rm = TRUE))
```

```{r}
cor.test(mean_PImd_by_family$mean_PImd, mean_piv_by_family$mean_PIv)

#Pearson's product-moment correlation
# 
# data:  mean_PImd_by_family$mean_PImd and mean_piv_by_family$mean_PIv
# t = 6.9007, df = 44, p-value = 1.599e-08
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  0.5446230 0.8362254
# sample estimates:
#       cor 
# 0.7209385

```

## **5. How is a paternal family's trait-specific plasticity as well as it's cross-trait (average) plasticity associated with breeding values for fitness?**

```{r}

HR22_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/HR22_bhat.Donor.mu.csv")
HR22_BV <- HR22_BV %>%
  rename(Donor = X) %>% 
  rename(BV_W = bhat.Donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))


HR23_BV<-read.csv("/Users/Helen/Desktop/nemo_field/aster/processed_data/HR23_bhat.Donor.mu.csv")
HR23_BV <- HR23_BV %>%
  rename(Donor = X) %>%   
  rename(BV_W = bhat.donor.mu) %>% 
  mutate(Donor = sub(".*Donor", "", Donor))

```

Merged the two datasets by Donor for 2022

```{r}
merged <- HR22_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated for 2022

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = 0.017918, df = 44, p-value = 0.9858
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2878225  0.2927697
# sample estimates:
#         cor 
# 0.002701258
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity for 2022

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- HR22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
   mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

#plot
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall Plasticity",
    y = "HR BV(W) 2022"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/HR_BV(W)2022_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

Merged the two datasets by Donor for 2023

```{r}
merged <- HR23_BV %>%
  select(Donor, BV_W) %>%       # keep only relevant columns
  inner_join(mean_piv_by_family %>% 
               select(Donor, mean_PIv),
             by = "Donor")
```

See if BV(W) and Plasticity are correlated in 2023

```{r}
cor_pearson <- cor(merged$BV_W, merged$mean_PIv, use = "complete.obs")
cor_pearson

cor.test(merged$BV_W, merged$mean_PIv)

#	Pearson's product-moment correlation
# 
# data:  merged$BV_W and merged$mean_PIv
# t = -1.9212, df = 44, p-value = 0.06121
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.52601066  0.01316897
# sample estimates:
#        cor 
# -0.2781926 
```

Make a scatterplot of Breeding value for Fitness and mean absolute plasticity in 2023

```{r}
library(dplyr)
library(ggplot2)

#Merge datasets by Donor
merged <- HR23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    mean_piv_by_family %>% select(Donor, mean_PIv),
    by = "Donor"
  )

#Scatterplot
p <- ggplot(merged, aes(x = mean_PIv, y = BV_W)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(
    x = "Overall plasticity",
    y = "HR BV(W) 2023"
  ) +
  theme_bw(base_size = 16) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    axis.title = element_text(face = "bold")
  )

print(p)

ggsave("Figures/HR_BV(W)2023_plasticity.jpeg", #save the plot
       plot = p, width = 8, height = 6, dpi = 700)
```

#### Determine if breeding values are correlated with any specific traits:

See if BV(W) in 2022 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- HR22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = 0.98654, df = 44, p-value = 0.3293
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.1495778  0.4194916
# sample estimates:
#      cor 
# 0.147108
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- HR22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)


# 	Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = 0.59632, df = 44, p-value = 0.554
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.2061176  0.3702132
# sample estimates:
#        cor 
# 0.08953802  
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- HR22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)


#Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -0.40086, df = 44, p-value = 0.6905
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3445858  0.2340750
# sample estimates:
#         cor 
# -0.06032197 

```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- HR22_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)


#	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -0.66727, df = 44, p-value = 0.5081
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3793654  0.1958999
# sample estimates:
#        cor 
# -0.1000902 
```

Determine if breeding values are correlated with any specific traits: See if BV(W) in 2023 and Plasticity in corolla diameter are correlated

```{r}
piv_df_corolla <- piv_df %>%
  filter(Trait == "mean_corolla_diam_mm_SEG")

merged_corolla <- HR23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_corolla %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_corolla$BV_W, merged_corolla$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_corolla$BV_W, merged_corolla$PIv)

# 	Pearson's product-moment correlation
# 
# data:  merged_corolla$BV_W and merged_corolla$PIv
# t = -2.0535, df = 44, p-value = 0.046
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.539692335 -0.005937915
# sample estimates:
#        cor 
# -0.2957262
```

See if BV(W) and Plasticity in SLA are correlated

```{r}
piv_df_SLA <- piv_df %>%
  filter(Trait == "mean_SLA_SEG")

merged_SLA <- HR23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_SLA %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_SLA$BV_W, merged_SLA$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_SLA$BV_W, merged_SLA$PIv)

#Pearson's product-moment correlation
# 
# data:  merged_SLA$BV_W and merged_SLA$PIv
# t = -0.61176, df = 44, p-value = 0.5438
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.3722109  0.2038980
# sample estimates:
#        cor 
# -0.0918359
```

See if BV(W) and Plasticity in d13C are correlated

```{r}
piv_df_d13C <- piv_df %>%
  filter(Trait == "mean_d13C_SEG")

merged_d13C <- HR23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_d13C %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_d13C$BV_W, merged_d13C$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_d13C$BV_W, merged_d13C$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_d13C$BV_W and merged_d13C$PIv
# t = -1.8585, df = 44, p-value = 0.0698
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.51940696  0.02225247
# sample estimates:
#        cor 
# -0.2697884 
```

See if BV(W) and Plasticity in msm are correlated

```{r}
piv_df_msm <- piv_df %>%
  filter(Trait == "mean_msm_all")

merged_msm <- HR23_BV %>%
  select(Donor, BV_W) %>% 
  inner_join(
    piv_df_msm %>% select(Donor, PIv),
    by = "Donor"
  )

cor_pearson <- cor(merged_msm$BV_W, merged_msm$PIv, use = "complete.obs")
cor_pearson

cor.test(merged_msm$BV_W, merged_msm$PIv)

#	Pearson's product-moment correlation
# 
# data:  merged_msm$BV_W and merged_msm$PIv
# t = -1.2755, df = 44, p-value = 0.2088
# alternative hypothesis: true correlation is not equal to 0
# 95 percent confidence interval:
#  -0.4542278  0.1073538
# sample estimates:
#        cor 
# -0.1888291 
```

Linear Regression to determine if a correlation exists: \####################################################### Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and absolute plasticity are correlated:

```{r}

model_linear  <- lm(BV_W ~ mean_PIv, data = merged)
model_quad    <- lm(BV_W ~ mean_PIv + I(mean_PIv^2), data = merged)

anova(model_linear, model_quad)

#Analysis of Variance Table
# 
# Model 1: BV_W ~ mean_PIv
# Model 2: BV_W ~ mean_PIv + I(mean_PIv^2)
#   Res.Df    RSS Df  Sum of Sq     F Pr(>F)
# 1     44 23.079                           
# 2     43 23.079  1 0.00026648 5e-04 0.9823

##the linear model is better than the quadratic

summary(model_linear)
#Call:
# lm(formula = BV_W ~ mean_PIv, data = merged)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -1.08929 -0.50429 -0.09806  0.39300  1.65211 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   3.2730     0.2749  11.906 2.36e-15 ***
# mean_PIv     -5.0437     2.6254  -1.921   0.0612 .  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.7242 on 44 degrees of freedom
# Multiple R-squared:  0.07739,	Adjusted R-squared:  0.05642 
# F-statistic: 3.691 on 1 and 44 DF,  p-value: 0.06121

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear...

shapiro.test(residuals(model_linear)) # normal enough
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in corolla diameter are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_corolla)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_corolla)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     44 22.827                           
# 2     43 22.822  1 0.0048591 0.0092 0.9242

summary(model_linear)

#Call:
# lm(formula = BV_W ~ PIv, data = merged_corolla)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -1.12925 -0.63845 -0.00587  0.46522  1.74637 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   3.0618     0.1711  17.896   <2e-16 ***
# PIv          -4.8887     2.3807  -2.053    0.046 *  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.7203 on 44 degrees of freedom
# Multiple R-squared:  0.08745,	Adjusted R-squared:  0.06671 
# F-statistic: 4.217 on 1 and 44 DF,  p-value: 0.046

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear..

shapiro.test(residuals(model_linear)) #non-normal
shapiro.test(residuals(model_quad)) #non-normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in d13C are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_d13C)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_d13C)

anova(model_linear, model_quad)
# Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     44 23.194                           
# 2     43 23.190  1 0.0038544 0.0071  0.933
summary(model_linear)
#Call:
# lm(formula = BV_W ~ PIv, data = merged_d13C)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.89008 -0.59861 -0.09624  0.41682  2.14961 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)    2.494      0.190  13.127   <2e-16 ***
# PIv          -24.715     13.298  -1.858   0.0698 .  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.726 on 44 degrees of freedom
# Multiple R-squared:  0.07279,	Adjusted R-squared:  0.05171 
# F-statistic: 3.454 on 1 and 44 DF,  p-value: 0.0698

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in msm are correlated:

```{r}

model_linear  <- lm(BV_W ~ PIv, data = merged_msm)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_msm)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F Pr(>F)
# 1     44 24.123                           
# 2     43 23.950  1   0.17317 0.3109   0.58

summary(model_linear)

# Call:
# lm(formula = BV_W ~ PIv, data = merged_msm)
# 
# Residuals:
#     Min      1Q  Median      3Q     Max 
# -1.0598 -0.6138 -0.1078  0.4531  1.7651 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   3.0319     0.2213  13.700   <2e-16 ***
# PIv          -1.1376     0.8919  -1.275    0.209    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.7404 on 44 degrees of freedom
# Multiple R-squared:  0.03566,	Adjusted R-squared:  0.01374 
# F-statistic: 1.627 on 1 and 44 DF,  p-value: 0.2088


plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```

Looking at a linear and quadratic model to determine if breeding value for fitness in 2023 and plasticity in SLA are correlated:

```{r}
library(car)
library(lmtest)


model_linear  <- lm(BV_W ~ PIv, data = merged_SLA)
model_quad    <- lm(BV_W ~ PIv + I(PIv^2), data = merged_SLA)

anova(model_linear, model_quad)
#Analysis of Variance Table
# 
# Model 1: BV_W ~ PIv
# Model 2: BV_W ~ PIv + I(PIv^2)
#   Res.Df    RSS Df Sum of Sq      F  Pr(>F)  
# 1     44 24.804                              
# 2     43 23.169  1     1.635 3.0345 0.08866 .
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

summary(model_linear)
# Call:
# lm(formula = BV_W ~ PIv, data = merged_SLA)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -0.96057 -0.63119 -0.07697  0.42011  1.97130 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)   2.9015     0.2184  13.288   <2e-16 ***
# PIv          -0.9168     1.4986  -0.612    0.544    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.7508 on 44 degrees of freedom
# Multiple R-squared:  0.008434,	Adjusted R-squared:  -0.0141 
# F-statistic: 0.3742 on 1 and 44 DF,  p-value: 0.5438

summary(model_quad)

# Call:
# lm(formula = BV_W ~ PIv + I(PIv^2), data = merged_SLA)
# 
# Residuals:
#      Min       1Q   Median       3Q      Max 
# -1.07266 -0.57474 -0.02669  0.33974  1.84746 
# 
# Coefficients:
#             Estimate Std. Error t value Pr(>|t|)    
# (Intercept)    2.549      0.294   8.672 5.42e-11 ***
# PIv            6.666      4.593   1.451   0.1539    
# I(PIv^2)     -28.272     16.230  -1.742   0.0887 .  
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Residual standard error: 0.734 on 43 degrees of freedom
# Multiple R-squared:  0.0738,	Adjusted R-squared:  0.03072 
# F-statistic: 1.713 on 2 and 43 DF,  p-value: 0.1924

plot(model_linear)
plot(model_quad)

#vif(model_linear)
vif(model_quad) #multicolinear

shapiro.test(residuals(model_linear)) #not normal
shapiro.test(residuals(model_quad)) #not normal

bptest(model_linear)
bptest(model_quad)

outlierTest(model_linear)
outlierTest(model_quad)
```